\documentclass[11pt,oneside]{report}

\usepackage[margin=0.9in,headheight=24pt]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage[protrusion=true,expansion=false]{microtype}

\usetikzlibrary{arrows.meta,calc,decorations.pathmorphing,positioning}

\definecolor{Ink}{HTML}{1B2430}
\definecolor{Muted}{HTML}{5B6472}
\definecolor{Accent}{HTML}{0077B6}
\definecolor{AccentTwo}{HTML}{D95F02}
\definecolor{Soft}{HTML}{F5F7FA}
\definecolor{SoftBlue}{HTML}{EAF4FB}
\definecolor{SoftOrange}{HTML}{FFF2E8}

\renewcommand{\familydefault}{\sfdefault}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist{itemsep=2pt,topsep=4pt}

\hypersetup{
  colorlinks=true,
  linkcolor=Accent,
  urlcolor=Accent,
  citecolor=Accent
}

\lstdefinestyle{pythonstyle}{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{Soft},
  columns=fullflexible,
  keepspaces=true,
  breaklines=true,
  frame=leftline,
  framerule=2pt,
  rulecolor=\color{Accent},
  xleftmargin=0.8em,
  framesep=0.6em,
  showstringspaces=false
}
\lstset{style=pythonstyle}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\scopebox}[2]{%
  \par\smallskip
  \noindent\fcolorbox{#1!55!black}{#1!10}{%
    \begin{minipage}{0.96\linewidth}
    #2
    \end{minipage}}
  \par\smallskip
}
\input{../docs/assets/mr_spin_dynamics_logo_tikz.tex}

\titleformat{\chapter}[block]
  {\huge\bfseries\color{Ink}}
  {\color{Accent}\thechapter}
  {0.75em}
  {}
\titlespacing*{\chapter}{0pt}{-0.15in}{1.1em}
\titleformat{\section}
  {\Large\bfseries\color{Ink}}
  {\thesection}
  {0.6em}
  {}
\titleformat{\subsection}
  {\large\bfseries\color{Ink}}
  {\thesubsection}
  {0.6em}
  {}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textcolor{Muted}{MRSpinDynamics Python}}
\rhead{\textcolor{Muted}{User Manual}}
\cfoot{\textcolor{Muted}{\thepage}}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\headrule}{\hbox to\headwidth{\color{Accent!35}\leaders\hrule height \headrulewidth\hfill}}

\begin{document}
\pagenumbering{gobble}
\begin{titlepage}
\pagecolor{Soft}
\color{Ink}
\vspace*{0.55in}
\begin{center}
\MRSpinDynamicsLogo[0.85]
\end{center}
\vspace{0.45in}
{\Huge\bfseries MRSpinDynamics\par}
\vspace{0.25in}
{\Large Python User Manual\par}
\vspace{0.15in}
{\large Python magnetic-resonance spin-dynamics workspace\par}
\vspace{0.45in}
{\color{Accent}\rule{0.82\linewidth}{3pt}\par}
\vspace{0.35in}
\begin{minipage}{0.78\linewidth}
\large
Simulation workflows for magnetic-resonance spin dynamics in possibly
non-uniform and time-varying fields, with validated MATLAB-compatible
spin-1/2 NMR Bloch kernels, Python-native analysis tools, scoped
scalar-coupled spin-1/2 models, pulsed NQR extensions, and first ESR/EPR
Hamiltonian and pulse-simulation helpers.
\end{minipage}
\vfill
{\large June 2026\par}
\end{titlepage}
\nopagecolor
\clearpage
\pagenumbering{roman}
\tableofcontents
\clearpage
\pagenumbering{arabic}

\chapter{Physical Scope and Assumptions}

The MATLAB-compatible workflow layer in PythonSpinDynamics models a bath of
uncoupled spin-1/2 nuclei evolving in a static, spatially non-uniform, or
time-varying main field \(B_0(\mathbf{r},t)\) and applied RF fields. The
elementary object in that layer is an isochromat: a subensemble of spin-1/2
nuclei that shares an offset, RF sensitivity, relaxation constants, and probe
response.

\scopebox{Accent}{
\textbf{Modeled directly in the core Bloch workflows.} Independent spin-1/2
magnetization vectors; offset distributions from \(B_0(\mathbf{r},t)\); RF
rotations; free precession; phenomenological \(T_1\) and \(T_2\) relaxation;
probe transmit and receive filtering; optional deterministic radiation damping;
optional motion through field maps.
}

\scopebox{Accent}{
\textbf{Scalar-coupling extension.} The \code{spin\_dynamics.coupling}
namespace adds small dense spin-1/2 systems with scalar J couplings, analytic
low-field J-editing models, ideal TANGO-B filter curves, and initial SLIC
simulations. These models are intentionally scoped and are documented in
Chapter~\ref{chap:jcoupling}.
}

\scopebox{Accent}{
\textbf{Pulsed NQR extension.} The \code{spin\_dynamics.nqr} namespace adds
dense single-site quadrupolar spin tools for selective pulsed NQR, including
spin-1 and spin-3/2 transition metadata, single-crystal and powder orientation
averages, SLSE echo trains, two-frequency population transfer, EFG
inhomogeneity, finite-window SLSE spectra, and weak-B0 perturbations. These
models are documented in Chapter~\ref{chap:nqr}.
}

\scopebox{Accent}{
\textbf{ESR/EPR extension.} The \code{spin\_dynamics.esr} namespace adds
single-electron spin-1/2 ESR tools, including scalar or anisotropic \(g\)
tensors, single-crystal and powder spectra, CW absorption/derivative
lineshapes, static \(g\)-strain and field-strain distributions, rectangular
pulsed ESR FID and Hahn-echo simulations with Liouville-space \(T_1/T_2\)
relaxation, and a first dense isotropic electron-nuclear hyperfine model.
These tools are documented in Chapter~\ref{chap:esr}.
}

\scopebox{AccentTwo}{
\textbf{Still outside the package scope.} Dipolar coupling networks, chemical
exchange, arbitrary nonselective multi-quantum pulse sequences, shaped sample
microstructure beyond the supplied field and density maps, microscopic spin
noise, and large sparse coupled-spin simulations. Spin-spin effects enter the
Bloch workflows only indirectly when represented by effective relaxation times
such as \(T_2\) or \(T_2^*\).
}

This scope is deliberate. The main package follows MATLAB's classical
Bloch/isochromat viewpoint, which is appropriate for many pulse, probe,
imaging, and relaxation studies where the spins can be treated as independent
subensembles. The coupled-spin namespace adds targeted quantum spin-1/2 tools
for scalar J-coupling phenomena without turning the full package into a general
quantum spin Hamiltonian framework.

\section{Spin Bath Cartoon}

\begin{center}
\begin{tikzpicture}[x=1cm,y=1cm,>=Latex]
  \fill[SoftBlue] (-0.4,-0.4) rectangle (9.7,3.8);
  \foreach \x/\y/\a in {
    0.5/0.6/25, 1.4/1.1/70, 2.1/0.55/130, 2.9/1.55/10,
    3.7/0.85/95, 4.4/1.45/45, 5.2/0.65/150, 6.0/1.2/80,
    6.8/0.55/20, 7.6/1.55/115, 8.5/0.85/60}
  {
    \draw[Accent,thick,->] (\x,\y) -- ++(\a:0.45);
    \fill[Ink] (\x,\y) circle (1.5pt);
  }
  \foreach \x in {0.7,1.7,...,8.7}
  {
    \draw[Muted!35,->] (\x,2.0) -- ++(0,1.1);
  }
  \node[Muted] at (8.4,3.45) {$B_0(\mathbf{r},t)$};
  \draw[AccentTwo,very thick,decorate,decoration={snake,amplitude=1.5pt,segment length=8pt},->]
    (-0.1,2.65) -- (3.0,2.65);
  \node[AccentTwo] at (1.35,3.05) {RF pulse};
  \draw[Ink!45,dashed] (3.2,0.15) -- (7.3,1.85);
  \node[Ink!55,align=center] at (5.75,2.15) {core workflows:\\no explicit links};
  \node[align=left,Ink] at (4.7,-0.05) {Independent isochromats sample the local field and RF sensitivity.};
\end{tikzpicture}
\end{center}

\section{Consequences for Users}

\begin{itemize}
  \item Use offset grids, field maps, or moving particles to represent
  spatially varying \(B_0\), transmit \(B_1\), and receive \(B_1\) effects.
  \item Use relaxation constants to represent ensemble dephasing and energy
  exchange that are outside the explicit isochromat model.
  \item Use \code{spin\_dynamics.coupling} when a small spin-1/2 scalar
  J-coupled model is the phenomenon of interest.
  \item Use \code{spin\_dynamics.esr} for single-electron ESR spectra, pulsed
  ESR FID/echo simulations, static disorder studies, and simple isotropic
  electron-nuclear hyperfine doublets.
  \item Use \code{spin\_dynamics.nqr} when selective pulsed NQR of a small
  quadrupolar site is the phenomenon of interest.
  \item Do not use the package for chemical exchange, large coupled-spin
  networks, or arbitrary nonselective multi-quantum coherence pathways.
\end{itemize}

\chapter{Overview}

PythonSpinDynamics is a Python port of the MATLAB spin-dynamics simulation
package. The MATLAB tree under \code{../MATLABSpinDynamics/SpinDynamicsUpdated/Version\_2/code}
remains the reference for numerical behavior, while this workspace provides a
typed, testable Python API for simulations, validation fixtures, examples, and
new analysis workflows.

The supported surface currently includes ideal, tuned, untuned, and matched
probe CPMG workflows; finite echo trains; FID; phase-encoded imaging; diffusion
CPMG; time-varying-field CPMG; WURST inversion and matched WURST-CPMG;
radiation damping; moving-isochromat motion helpers; received-signal noise;
inverse-Laplace analysis; compact pulse-optimization scaffolds; and small
spin-1/2 scalar-coupling tools for low-field J-editing, TANGO-B filtering, and
SLIC-style spin-lock response; and early pulsed NQR helpers for quadrupolar
sites, powder averaging, SLSE, two-frequency population transfer, EFG
inhomogeneity, spin-3/2 transition metadata, and weak-static-field spectra.

\section{Package Layout}

\begin{longtable}{p{0.36\linewidth}p{0.56\linewidth}}
\toprule
Path & Purpose \\
\midrule
\code{src/spin\_dynamics/} & Installable Python package. \\
\code{examples/} & CLI examples and plotting scripts. \\
\code{tests/} & Smoke tests, validation tests, and unit tests. \\
\code{validation/fixtures/} & MATLAB/Octave fixture arrays. \\
\code{validation/octave/} & Fixture generation scripts. \\
\code{docs/python\_api/} & Lightweight Markdown API notes. \\
\code{docs/validation\_results.md} & Current validation status and tolerance notes. \\
\code{docs/radiation\_damping.md} & Focused radiation-damping model notes. \\
\bottomrule
\end{longtable}

\chapter{Installation and Test Runs}

\section{Editable Install}

PythonSpinDynamics requires Python 3.10 or newer. The core install depends on
NumPy 1.24 or newer and does not require MATLAB at runtime. MATLAB is only
needed to regenerate the complete MATLAB reference fixture set; Octave can
regenerate the smaller dependency-light fixtures.

From \code{PythonSpinDynamics}, install the package into an active environment:

\begin{lstlisting}[language=bash]
python -m pip install -e .
\end{lstlisting}

Optional extras are provided for development, plotting, benchmarking, and
SciPy-backed optimization:

\begin{lstlisting}[language=bash]
python -m pip install -e ".[dev,opt,plot,bench]"
\end{lstlisting}

The core package depends only on NumPy. The \code{opt} extra enables SciPy-backed
continuous optimizers, non-negative inverse-Laplace solves, and MATLAB
\code{.mat} result import/export helpers. The \code{plot} extra enables the
Matplotlib examples.

\section{Tests}

Use the smoke tier during normal editing:

\begin{lstlisting}[language=bash]
python -m unittest tests.smoke_tests
\end{lstlisting}

Run the full validation suite before committing numerical or workflow changes:

\begin{lstlisting}[language=bash]
python -m unittest discover -s tests
\end{lstlisting}

\chapter{Model Conventions}

\section{Coherence Ordering}

Low-level kernels preserve MATLAB's coherence ordering:
\[
  \mathbf{M} =
  \begin{bmatrix}
    M_0 & M_- & M_+
  \end{bmatrix}^{\mathsf{T}} .
\]
This ordering differs from many Bloch-vector presentations, so new kernels and
validation tests should state the expected ordering explicitly.

\section{Normalized Time}

The core MATLAB routines usually normalize time to the nominal RF amplitude
\( \omega_1 = 1 \). In these units, a 90-degree hard pulse has duration
\[
  t_{90} = \frac{\pi}{2},
\]
and a 180-degree hard pulse has duration
\[
  t_{180} = \pi .
\]
Workflow helpers that expose physical seconds convert into these normalized
units before calling the lower-level kernels.

\section{Pulse Timing Cartoon}

A minimal spin-echo or CPMG building block is a sequence of RF pulses, free
precession windows, and acquisition windows. The package represents these as
ordered intervals with pulse matrices and free-precession delays.

\begin{center}
\begin{tikzpicture}[x=1cm,y=1cm,>=Latex]
  \draw[Ink,->] (0,0) -- (10.5,0) node[right] {$t$};
  \draw[Accent,fill=Accent!18,rounded corners=1pt] (0.7,0) rectangle (1.1,1.5);
  \node[Accent] at (0.9,1.85) {$90_x$};
  \draw[Accent,fill=Accent!18,rounded corners=1pt] (4.4,0) rectangle (5.0,2.2);
  \node[Accent] at (4.7,2.55) {$180_y$};
  \draw[Accent,fill=Accent!18,rounded corners=1pt] (7.9,0) rectangle (8.5,2.2);
  \node[Accent] at (8.2,2.55) {$180_y$};
  \draw[Muted,<->] (1.15,-0.35) -- (4.35,-0.35);
  \node[Muted] at (2.75,-0.68) {$\tau$};
  \draw[Muted,<->] (5.05,-0.35) -- (7.85,-0.35);
  \node[Muted] at (6.45,-0.68) {$2\tau$};
  \draw[AccentTwo,thick] (2.55,0.15) .. controls (3.0,0.8) and (3.45,0.8) .. (3.9,0.15);
  \node[AccentTwo] at (3.25,1.05) {echo};
  \draw[AccentTwo,thick] (6.25,0.15) .. controls (6.7,1.0) and (7.15,1.0) .. (7.6,0.15);
  \node[AccentTwo] at (6.95,1.25) {echo};
  \draw[Ink!35,fill=Ink!8] (2.35,-0.02) rectangle (4.05,-0.18);
  \draw[Ink!35,fill=Ink!8] (6.05,-0.02) rectangle (7.75,-0.18);
\end{tikzpicture}
\end{center}

\section{Effective Rotations}

For an isochromat with normalized offset \(\Delta \omega\), RF amplitude
\(\omega_1\), phase \(\phi\), and pulse duration \(t_p\), the rotating-frame
effective field is
\[
  \boldsymbol{\Omega} =
  \begin{bmatrix}
    \omega_1 \cos \phi \\
    \omega_1 \sin \phi \\
    \Delta \omega
  \end{bmatrix},
  \qquad
  \Omega = \sqrt{\omega_1^2 + \Delta\omega^2}.
\]
The magnetization update is a rotation by angle \(\Omega t_p\) about
\(\boldsymbol{\Omega}/\Omega\). The Python rotation helpers expose this
calculation through \code{spin\_dynamics.core.rotations} and return array shapes
chosen to match the MATLAB reference fixtures.

\section{Rotation Cartoon}

In the rotating frame, a rectangular RF segment rotates the magnetization about
an effective axis set by RF phase, RF amplitude, and off-resonance offset.

\begin{center}
\begin{tikzpicture}[x=1cm,y=1cm,>=Latex]
  \coordinate (O) at (0,0);
  \draw[Ink!55,->] (O) -- (3.0,0) node[right] {$x$};
  \draw[Ink!55,->] (O) -- (0,2.7) node[above] {$z$};
  \draw[Ink!55,->] (O) -- (-1.5,-1.2) node[left] {$y$};
  \draw[AccentTwo,very thick,->] (O) -- (0.75,2.05) node[above right] {$\boldsymbol{\Omega}$};
  \draw[Accent,very thick,->] (O) -- (2.15,0.55) node[right] {$\mathbf{M}_{\mathrm{before}}$};
  \draw[Accent,very thick,->] (O) -- (0.65,1.85) node[left] {$\mathbf{M}_{\mathrm{after}}$};
  \draw[Accent!70,thick,->] (1.55,0.42) arc[start angle=12,end angle=68,radius=1.6];
  \node[Muted,align=left] at (5.3,1.6) {
    rotation angle: $\Omega t_p$\\
    RF phase sets transverse axis\\
    offset tilts the axis toward $z$
  };
  \draw[AccentTwo!55,dashed] (0.75,2.05) -- (0.75,0);
  \draw[AccentTwo!55,dashed] (0.75,0) -- (0,0);
\end{tikzpicture}
\end{center}

\section{Relaxation}

Free-precession intervals with relaxation use the usual exponential factors:
\[
  M_{xy}(t+\Delta t) =
    M_{xy}(t)\exp\left(-\frac{\Delta t}{T_2}\right)
    \exp(-\ii \Delta\omega\,\Delta t),
\]
\[
  M_z(t+\Delta t) =
    M_{\mathrm{th}} +
    \left(M_z(t)-M_{\mathrm{th}}\right)
    \exp\left(-\frac{\Delta t}{T_1}\right).
\]
Finite acquisition workflows assemble these intervals from pulse-sequence
parameters and then call the \code{arb10}-style kernels.

\section{Echo Conversion}

The echo utilities direct-sum a received spectrum \(S(\Delta\omega)\) over the
offset grid:
\[
  e(t) =
    \sum_k S(\Delta\omega_k)
    \exp(\ii \Delta\omega_k t)\,\Delta\omega .
\]
The implementation uses the same direct-sum convention as the MATLAB
\code{calc\_time\_domain\_echo} reference path. FID traces use the corresponding
FID conversion helper.

\section{Radiation Damping}

The deterministic radiation-damping back-action model follows the local
Measurements text convention:
\[
  T_{\mathrm{rd}} =
    \frac{2}{\gamma \mu_0 \eta M_0 Q},
\]
where \(\gamma\) is the gyromagnetic ratio, \(\eta\) is the magnetic-energy fill
factor, \(M_0\) is the equilibrium magnetization density, and \(Q\) is the
loaded probe quality factor.

The analytic on-resonance hard-pulse FID envelope used by the test suite is
\[
  |M_{xy}(t)| =
    \frac{M_0}
    {\cosh\left(t/T_{\mathrm{rd}} - \log \tan(\theta/2)\right)}.
\]
The finite-sequence implementation can add feedback during free-precession and
acquisition windows, and can optionally apply an operator-split approximation
during RF pulse matrices.

\chapter{Scalar J-Coupling Models}
\label{chap:jcoupling}

Scalar J coupling is field independent, so it can remain visible in weak or
grossly inhomogeneous magnetic fields where chemical-shift resolution is
compressed. The \code{spin\_dynamics.coupling} namespace adds a scoped set of
spin-1/2 tools for these cases. It is separate from the MATLAB-compatible Bloch
workflow layer: Bloch workflows propagate independent isochromats, while the
coupling layer propagates small dense Hilbert-space systems or evaluates
closed-form J-editing response curves.

\scopebox{Accent}{
\textbf{Use this layer for:} heteronuclear low-field J-editing curves,
known-J spectrum fitting, ideal TANGO-B filter selectivity, and exploratory
two-spin or few-spin scalar-coupled Hamiltonian simulations.
}

\scopebox{AccentTwo}{
\textbf{Current limits:} spin-1/2 only; dense matrices only; no relaxation
superoperators; no chemical exchange; no quadrupolar nuclei; no large sparse
coupled-spin networks; no full pulse-sequence compiler for arbitrary
multi-quantum pathways.
}

\section{J-Coupled Spin Cartoon}

\begin{center}
\begin{tikzpicture}[x=1cm,y=1cm,>=Latex]
  \fill[SoftBlue] (-0.5,-0.45) rectangle (9.8,3.7);
  \draw[Muted!35,->] (1.0,0.25) -- (1.0,2.9);
  \draw[Muted!35,->] (3.0,0.25) -- (3.0,2.9);
  \node[Muted] at (2.0,3.25) {$B_0(\mathbf{r},t)$ may be weak or non-uniform};
  \fill[Accent!18] (1.0,1.35) circle (0.38);
  \fill[AccentTwo!18] (3.0,1.35) circle (0.38);
  \draw[Accent,very thick,->] (1.0,1.35) -- ++(70:0.75);
  \draw[AccentTwo,very thick,->] (3.0,1.35) -- ++(120:0.75);
  \node[Ink] at (1.0,0.75) {A};
  \node[Ink] at (3.0,0.75) {B};
  \draw[Ink!55,thick,decorate,decoration={snake,amplitude=1.3pt,segment length=6pt}]
    (1.42,1.35) -- (2.58,1.35);
  \node[Ink] at (2.0,1.75) {$J_{AB}$};
  \draw[Ink,->] (4.5,0.75) -- (9.1,0.75) node[right] {$\tau$};
  \draw[Accent,thick] (4.7,1.45) cos (5.55,0.75) sin (6.4,0.05)
    cos (7.25,0.75) sin (8.1,1.45);
  \node[Accent,align=center] at (6.4,2.2) {echo amplitude\\oscillates with \(J\)};
  \node[Ink,align=left] at (6.95,-0.05) {coupling frequency is not set by \(B_0\)};
\end{tikzpicture}
\end{center}

\section{Dense Spin-1/2 Hamiltonians}

A coupled system is built from offsets \(\nu_i\) in hertz and symmetric scalar
couplings \(J_{ij}\) in hertz:

\begin{lstlisting}[language=Python]
from spin_dynamics.coupling import (
    coupled_spin_system,
    isotropic_j_hamiltonian,
    zeeman_hamiltonian,
)

system = coupled_spin_system(
    offsets_hz=[-0.35, 0.35],
    couplings_hz=[[0.0, 7.0], [7.0, 0.0]],
    labels=["A", "B"],
)
hamiltonian = zeeman_hamiltonian(system) + isotropic_j_hamiltonian(system)
\end{lstlisting}

Hamiltonian builders return angular-frequency matrices in radians per second
and use spin-1/2 operators with eigenvalues \(\pm 1/2\). The rotating-frame
offset Hamiltonian is
\[
  H_Z = 2\pi \sum_i \nu_i I_{iz}.
\]
For weakly coupled systems, the secular scalar Hamiltonian is
\[
  H_J^{\mathrm{sec}} =
    2\pi \sum_{i<j} J_{ij} I_{iz} I_{jz},
\]
while the isotropic scalar Hamiltonian is
\[
  H_J^{\mathrm{iso}} =
    2\pi \sum_{i<j} J_{ij}
      \left(I_{ix}I_{jx}+I_{iy}I_{jy}+I_{iz}I_{jz}\right).
\]
The dense propagator evaluates
\[
  \rho(t) = U(t)\rho(0)U^\dagger(t),
  \qquad
  U(t)=\exp\left(-\ii Ht\right),
\]
and is intended for small systems where dense matrices are acceptable.

\section{Inhomogeneous B0/B1 Isochromats}

The bridge between the coupled-spin layer and the Bloch workflow layer is a
coupled isochromat ensemble. Each isochromat contains the same scalar-coupled
spin network, but samples a different local field and receive weight. For
isochromat \(k\), the local spin offsets are
\[
  \nu_{i,k} = \nu_i^{(0)} + s_i\,\Delta\nu_k ,
\]
where \(\nu_i^{(0)}\) is the base spin offset, \(\Delta\nu_k\) is the local
B0 offset in hertz, and \(s_i\) is a per-spin offset scale. The default
\(s_i=1\) adds the same field offset to every spin. Heteronuclear or
carrier-specific models can supply different scales.

During an RF or spin-lock interval, the nominal nutation frequency is scaled by
the local transmit field:
\[
  \omega_{1,k} = b_{1,k}^{\mathrm{tx}} \omega_1 .
\]
The ensemble signal is accumulated as
\[
  S = \sum_k w_k b_{1,k}^{\mathrm{rx}}
      \operatorname{Tr}\!\left(\rho_k D\right),
\]
where \(w_k\) is the isochromat weight, \(b_{1,k}^{\mathrm{rx}}\) is the local
receive scale, and \(D\) is the detection operator.

\begin{lstlisting}[language=Python]
from spin_dynamics.coupling import (
    coupled_isochromat_ensemble,
    free_precession_step,
    rf_step,
    simulate_coupled_isochromat_sequence,
)

ensemble = coupled_isochromat_ensemble(
    system,
    b0_offsets_hz=[-2.0, 0.0, 2.0],
    weights=[0.25, 0.5, 0.25],
    b1_tx_scale=[0.9, 1.0, 1.1],
    b1_rx_scale=[1.0, 1.0, 0.8],
)

result = simulate_coupled_isochromat_sequence(
    ensemble,
    [
        rf_step(duration=0.25, nutation_hz=1.0, phase=1.5708),
        free_precession_step(duration=0.01),
    ],
    initial_axis="x",
    detect_axis="x",
)
\end{lstlisting}

For time-varying fields, each sequence step may override the ensemble B0
offsets or B1 transmit scale. This keeps the first implementation close to the
existing isochromat approach: the coupled-spin Hilbert space is dense and small,
while field inhomogeneity is represented by an outer ensemble sum.

\section{Low-Field J-Editing}

Analytic J-editing helpers model response curves as sums of field-independent
cosine modulations:
\[
  s(\tau) =
    b + \sum_m a_m
      \cos^{p_m}\!\left(2\pi N J_m \tau\right),
\]
where \(\tau\) is the encoding time, \(N\) is the number of encoding cycles,
\(J_m\) is a known or hypothesized coupling in hertz, \(a_m\) is its amplitude,
and \(p_m\) is an optional multiplicity exponent. Proton-detected models use
\(p_m=1\). Carbon-detected \(\mathrm{CH}_n\) groups use \(p_m=n\).

\begin{lstlisting}[language=Python]
from spin_dynamics.coupling import (
    carbon_detected_j_modulation,
    fit_known_j_spectrum,
    j_modulation_curve,
)

signal = j_modulation_curve(
    encoding_times,
    couplings_hz=[125.0, 160.0],
    amplitudes=[0.85, 0.15],
)

carbon_signal = carbon_detected_j_modulation(
    encoding_times,
    couplings_hz=[125.0, 160.0],
    abundances=[0.85, 0.15],
    proton_counts=[2, 1],
)

fit = fit_known_j_spectrum(
    encoding_times,
    signal,
    couplings_hz=[125.0, 160.0],
    include_background=False,
)
\end{lstlisting}

\code{fit\_known\_j\_spectrum} solves a linear least-squares problem once the
candidate \(J_m\) values are supplied. It returns amplitudes, optional
background, fitted signal, residual, rank, and residual norm.

\section{TANGO-B Filtering}

The ideal TANGO-B helper evaluates the coupled-spin transverse filter envelope
\[
  A(J;t_d) = \sin^2\!\left(\pi J t_d\right).
\]
When a target coupling \(J_\star\) and positive odd order \(n\) are supplied,
the delay is selected as
\[
  t_d = \frac{n}{2J_\star}.
\]

\begin{lstlisting}[language=Python]
from spin_dynamics.coupling import tango_b_filter

response = tango_b_filter(
    couplings_hz,
    target_coupling_hz=160.0,
    order=3,
)
\end{lstlisting}

This model is idealized: it reports the scalar-coupling selectivity envelope,
not RF imperfections, relaxation, diffusion, or probe filtering.

\section{SLIC Response}

Spin-lock induced crossing (SLIC) is modeled by adding an RF spin-lock
Hamiltonian to the static coupled-spin Hamiltonian:
\[
  H_{\mathrm{SLIC}} =
    H_Z + H_J^{\mathrm{iso}}
    + 2\pi \omega_1 \sum_i I_{ix}.
\]
The helper scans spin-lock nutation frequencies \(\omega_1\) in hertz and
returns the remaining transverse magnetization and its corresponding dip:

\begin{lstlisting}[language=Python]
from spin_dynamics.coupling import (
    coupled_spin_system,
    simulate_slic_spectrum,
    two_spin_slic_transfer_time,
)

system = coupled_spin_system(
    offsets_hz=[-0.35, 0.35],
    couplings_hz=[[0.0, 7.0], [7.0, 0.0]],
)
duration = two_spin_slic_transfer_time(offset_difference_hz=0.7)
result = simulate_slic_spectrum(
    system,
    nutation_frequencies_hz,
    spin_lock_time=duration,
)
\end{lstlisting}

For a nearly equivalent two-spin system, the SLIC dip is expected near the
coupling frequency. The helper is exploratory and dense; broad homonuclear
networks will need sparse or specialized methods.

\chapter{Electron Spin Resonance Models}
\label{chap:esr}

The \code{spin\_dynamics.esr} namespace adds an ESR/EPR extension for
single-electron spin-1/2 systems. The first layer is intentionally compact: it
uses dense matrices, explicit \(g\)-tensor axes, and small product Hilbert
spaces for hyperfine examples. Hamiltonians are expressed in radians per
second, while public resonance frequencies, linewidths, and hyperfine
constants are expressed in hertz.

\scopebox{Accent}{
\textbf{Use this layer for:} single-electron ESR resonance calculations,
anisotropic \(g\)-tensor single-crystal and powder spectra, CW absorption or
first-derivative display, static \(g\)-strain and field-strain broadening,
rectangular pulsed ESR FID and Hahn-echo simulations, Liouville-space
\(T_1/T_2\) relaxation, and first electron-nuclear isotropic hyperfine
doublets.
}

\scopebox{AccentTwo}{
\textbf{Current limits:} electron spin \(S=1/2\) only in the main ESR system;
dense matrices only; hyperfine couplings are isotropic \(A\,\mathbf{S}\cdot
\mathbf{I}\); no anisotropic \(A\) tensors, exchange, electron-electron dipolar
coupling, higher-spin zero-field splitting, saturation, or microwave resonator
model yet.
}

\section{Zeeman Hamiltonian and \(g\) Tensors}

The single-electron Zeeman Hamiltonian is
\[
  H_Z =
    2\pi \frac{\mu_B}{h}\,
    \mathbf{B}_0^{\mathsf{T}}\, \mathbf{g}\, \mathbf{S},
\]
where \(\mu_B/h\) is exposed as \code{BOHR\_MAGNETON\_HZ\_PER\_T}. The \(g\)
tensor may be supplied as a scalar, a three-component principal-axis vector, or
a full \(3\times 3\) matrix. For a static-field direction \(\hat{\mathbf{n}}\),
the effective \(g\) value is
\[
  g_{\mathrm{eff}}(\hat{\mathbf{n}})
    = \left|\mathbf{g}^{\mathsf{T}}\hat{\mathbf{n}}\right|,
\]
and the spin-1/2 transition frequency is
\[
  \nu_{\mathrm{ESR}} =
    \frac{\mu_B}{h}\,B_0\,g_{\mathrm{eff}} .
\]

\begin{lstlisting}[language=Python]
from spin_dynamics.esr import (
    ESRSpinSystem,
    effective_g_value,
    resonance_field_tesla,
    resonance_frequency_hz,
)

system = ESRSpinSystem(g_tensor=[2.00, 2.08, 2.24])
g_z = effective_g_value(system, [0.0, 0.0, 1.0])
nu = resonance_frequency_hz(system, [0.0, 0.0, 0.34])
b_res = resonance_field_tesla(system, microwave_frequency_hz=9.5e9)
\end{lstlisting}

\section{CW Spectra, Lineshapes, and Disorder}

Frequency-swept and field-swept helpers broaden orientation-resolved ESR lines
with Gaussian or Lorentzian profiles. The same helpers can return absorption
or first-derivative CW-style spectra:

\begin{lstlisting}[language=Python]
from spin_dynamics.esr import simulate_field_sweep

result = simulate_field_sweep(
    system,
    microwave_frequency_hz=9.5e9,
    orientations="powder",
    broadening_tesla=0.2e-3,
    lineshape="lorentzian",
    detection_mode="derivative",
)
\end{lstlisting}

Static disorder is represented by weighted samples. The convenience
\code{static\_disorder\_grid} builds Gaussian quadrature-like samples for
diagonal \(g\)-strain and applied-field offsets:

\begin{lstlisting}[language=Python]
from spin_dynamics.esr import (
    simulate_field_sweep_distribution,
    static_disorder_grid,
)

samples = static_disorder_grid(
    system,
    g_std=[0.0, 0.0, 0.005],
    field_std_tesla=0.1e-3,
    g_points=5,
    field_points=5,
)

strained = simulate_field_sweep_distribution(
    samples,
    microwave_frequency_hz=9.5e9,
    orientations="powder",
    broadening_tesla=0.05e-3,
)
\end{lstlisting}

\section{Pulsed ESR}

The pulsed ESR helpers use a single rotating frame and a rotating-wave
approximation. The \code{nutation\_hz} parameter is the on-resonance spin-1/2
Rabi rate for the selected microwave \(B_1\) direction, so a rectangular pulse
of flip angle \(\theta\) has duration
\[
  t_p = \frac{\theta}{2\pi \nu_1}.
\]
In particular, a 90-degree pulse has duration \(1/(4\nu_1)\).

\begin{lstlisting}[language=Python]
import numpy as np

from spin_dynamics.esr import (
    flip_angle_duration,
    simulate_fid,
    simulate_hahn_echo,
)

carrier = resonance_frequency_hz(system, [0.0, 0.0, 0.34])
nu1 = 5e6
t90 = flip_angle_duration(np.pi / 2.0, nu1)
t180 = flip_angle_duration(np.pi, nu1)

fid = simulate_fid(
    system,
    [0.0, 0.0, 0.34],
    nutation_hz=nu1,
    pulse_duration_seconds=t90,
    times_seconds=np.linspace(0.0, 5e-6, 256),
    rf_frequency_hz=carrier,
)

echo = simulate_hahn_echo(
    system,
    [0.0, 0.0, 0.34],
    nutation_hz=nu1,
    excitation_duration_seconds=t90,
    refocus_duration_seconds=t180,
    tau_seconds=2e-6,
    times_seconds=np.linspace(0.0, 4e-6, 256),
    rf_frequency_hz=carrier,
)
\end{lstlisting}

\section{Pulsed Relaxation}

For physical \(T_1/T_2\) relaxation during pulses, free evolution, and
acquisition, pass an \code{ESRRelaxationModel}. The model acts in
Liouville space on the trace-zero high-temperature density-matrix deviation:
\(T_1\) damps population differences while preserving trace, and \(T_2\) damps
coherences.

\begin{lstlisting}[language=Python]
from spin_dynamics.esr import ESRRelaxationModel

relaxed = simulate_fid(
    system,
    [0.0, 0.0, 0.34],
    nutation_hz=nu1,
    pulse_duration_seconds=t90,
    times_seconds=np.linspace(0.0, 5e-6, 256),
    rf_frequency_hz=carrier,
    relaxation=ESRRelaxationModel(t1_seconds=50e-6, t2_seconds=3e-6),
)
\end{lstlisting}

The older \code{t2\_seconds} argument in the pulsed helpers is a simple
post-propagation envelope retained for quick examples. When using
\code{ESRRelaxationModel}, leave \code{t2\_seconds=inf} to avoid double-counting
coherence decay.

\section{Isotropic Hyperfine Coupling}

The first hyperfine layer models one electron spin-1/2 coupled isotropically to
one or more nuclei:
\[
  H =
    H_{eZ} + H_{nZ}
    + 2\pi\sum_k A_k\,\mathbf{S}\cdot\mathbf{I}_k ,
\]
with \(A_k\) in hertz. The helper scans the applied field, diagonalizes the
full dense electron-nuclear Hamiltonian at each point, and sums ESR-active
transitions according to the microwave \(B_1\) matrix element.

\begin{lstlisting}[language=Python]
from spin_dynamics.esr import (
    NuclearSite,
    electron_nuclear_system,
    simulate_hyperfine_field_sweep,
)

hf_system = electron_nuclear_system(
    [20e6],
    nuclei=[NuclearSite("H1", isotope="1H", gamma_hz_per_t=42.577e6)],
    g_tensor=2.0023,
)

hf_sweep = simulate_hyperfine_field_sweep(
    hf_system,
    microwave_frequency_hz=9.5e9,
    broadening_hz=0.5e6,
)
\end{lstlisting}

For one spin-1/2 nucleus in the high-field limit, the ESR line splits by
approximately \(A\) in frequency, or by
\[
  \Delta B \approx \frac{A}{(\mu_B/h)g}
\]
in a field-swept spectrum.

\chapter{Pulsed NQR Models}
\label{chap:nqr}

Nuclear quadrupole resonance is driven by the interaction between a nuclear
quadrupole moment and the local electric-field-gradient (EFG) tensor. The
\code{spin\_dynamics.nqr} namespace adds a scoped quadrupolar extension for
solid-state pulsed NQR experiments. It is intentionally separate from the
spin-1/2 Bloch and scalar-coupling layers.

\scopebox{Accent}{
\textbf{Use this layer for:} reduced two-level (selective) pulsed NQR of spin-1
sites; full \((2I+1)\)-level density-matrix FID and echo dynamics required for
spin-3/2; automatic reduced-vs-full model selection; spin-1 and spin-3/2
transition metadata; single-crystal and powder averages over local EFG
orientations; SLSE echo trains; two-frequency population transfer; static EFG
inhomogeneity; finite-window SLSE spectra; and weak-B0 line splitting.
}

\scopebox{AccentTwo}{
\textbf{Current limits:} dense single-site matrices only; two pulse models (the
reduced spin-1 selective pulse and the full \((2I+1)\)-level density-matrix
pulse). The full-model pulse uses a single-carrier rotating-wave approximation
addressing one transition band, so it covers spin-3/2 zero-field and
weak-Zeeman dynamics but is not yet a general multi-band higher-spin solver. The
reduced SLSE and population-transfer workflows remain spin-1; relaxation is
phenomenological rather than microscopic Redfield/dipolar; no multi-site
coupling between quadrupolar nuclei.
}

\section{Quadrupolar Site Model}

A quadrupolar site is described in the EFG principal-axis system. The zero-field
Hamiltonian used by the Python helper is
\[
  H_Q = 2\pi \nu_{\mathrm{scale}}
  \left[
    3 I_z^2 - I(I+1)\mathbf{1}
    + \eta\left(I_x^2 - I_y^2\right)
  \right],
\]
where \(I\) is the spin quantum number, \(\nu_Q\) is the quadrupole-frequency
parameter in hertz, and \(0 \le \eta \le 1\) is the EFG asymmetry parameter.
Hamiltonians are represented in radians per second. For spin-1, the package
uses \(\nu_{\mathrm{scale}}=\nu_Q/3\), matching the nitrogen-14 convention
where the zero-field lines are \(\nu_x=\nu_Q(1+\eta/3)\),
\(\nu_y=\nu_Q(1-\eta/3)\), and \(\nu_z=2\eta\nu_Q/3\). For spin-3/2,
\(\nu_{\mathrm{scale}}=\nu_Q/6\), so the zero-field chlorine-style NQR line is
\[
  \nu_{\mathrm{NQR}} = \nu_Q\sqrt{1+\eta^2/3}.
\]
Zero-frequency Kramers-doublet transitions are omitted from the transition
inventory.

Weak Zeeman perturbations can be added as
\[
  H_Z = -2\pi \gamma\, \mathbf{B}_0\cdot\mathbf{I},
\]
where \(\mathbf{B}_0\) is expressed in the EFG principal-axis system for the
current crystallite orientation. The default is \(B_0=0\), which is the usual
NQR case.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import QuadrupolarSite, diagonalize_site

site = QuadrupolarSite(
    spin=1,
    isotope="14N",
    quadrupole_frequency_hz=900e3,
    eta=0.3,
)

eigensystem = diagonalize_site(site)
for transition in eigensystem.transitions:
    print(transition.label, transition.frequency_hz)
\end{lstlisting}

For spin-1 at zero field, transitions are labeled by their dominant RF
polarization in the EFG principal-axis system: \code{x}, \code{y}, and
\code{z}. For the example above the line positions are approximately
\(990\), \(810\), and \(180\) kHz.

Spin-3/2 transition metadata can be inspected with the same diagonalization
helper:

\begin{lstlisting}[language=Python]
chlorine = QuadrupolarSite(
    spin=1.5,
    isotope="35Cl",
    quadrupole_frequency_hz=30e6,
    eta=0.1,
    gamma_hz_per_t=4.171e6,
)

for transition in diagonalize_site(chlorine).transitions:
    print(transition.frequency_hz)
\end{lstlisting}

\section{Two Modeling Regimes}
\label{sec:nqr-regimes}

Before propagating a pulse, the package chooses between two models. The choice
is governed by how many states the RF pulse can connect, not by the number of
spectral lines. The detailed reasoning is given in the technical note
\code{References/Pulsed\_NQR\_Spin\_Dynamics\_Narrative\_Rewrite}.

\begin{itemize}
\item \textbf{Reduced two-level model} (\code{SelectivePulse},
\code{simulate\_slse}): one pulse addresses one transition as an embedded
fictitious spin-1/2 rotation. This is honest only when the selected transition
is isolated. For a target transition \(t\), with
\(\Delta_{\mathrm{iso}}=\min_{j\neq t}|\omega_t-\omega_j|\) the spacing to the
nearest RF-active competitor, the validity condition is
\[
  \Delta_{\mathrm{iso}} \gg
  \max\!\left(\Omega_1,\ \Delta\omega_{\mathrm{pulse}},\ \Gamma\right),
\]
where \(\Omega_1\) is the effective nutation rate, \(\Delta\omega_{\mathrm{pulse}}\sim 1/t_p\)
is the pulse bandwidth, and \(\Gamma\) is the linewidth. A nonzero \(\eta\) is
necessary but \emph{not} sufficient: at \(\eta=0\) the spin-1 \(x\) and \(y\)
lines coincide and the reduction fails. The reduced path is restricted to
spin-1 for this reason.
\item \textbf{Full density-matrix model}
(\code{spin\_dynamics.nqr.full\_dynamics}): the complete \((2I+1)\)-level
density matrix is propagated. This is the correct general model and is required
for spin-3/2, whose single zero-field line connects two Kramers doublets---four
states hidden behind one frequency.
\end{itemize}

\section{Selective Pulse Approximation}

Most practical spin-1 pulsed NQR RF pulses are narrowband relative to the
separation between quadrupolar transitions. The reduced model uses this
selective limit: one pulse addresses one transition and acts as an embedded
two-level rotation inside the full \((2I+1)\)-level Hilbert space.

For a transition \(|a\rangle \leftrightarrow |b\rangle\), the transition dipole
vector is
\[
  \mathbf{d}_{ab} =
  \left(
    \langle a|I_x|b\rangle,
    \langle a|I_y|b\rangle,
    \langle a|I_z|b\rangle
  \right).
\]
For a local RF direction \(\hat{\mathbf{b}}_1\), the complex drive projection
is proportional to \(\hat{\mathbf{b}}_1\cdot\mathbf{d}_{ab}\). Preserving this
signed complex projection is essential for powder averages because transmit
and receive phases must pair consistently across crystallites.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import SelectivePulse, apply_selective_pulse

pulse = SelectivePulse(
    "x",
    duration_seconds=50e-6,
    nutation_hz=10e3,
)
\end{lstlisting}

The pulse helper also supports an optional RF frequency. If omitted, the RF is
placed on the selected transition. If supplied, the difference between the RF
frequency and transition frequency is treated as a two-level detuning.

The selective-pulse propagation routines support spin-1 sites only. Spin-3/2
nuclei such as \({}^{35}\mathrm{Cl}\) and \({}^{37}\mathrm{Cl}\) have degenerate
doublets at zero field, so their pulsed response is handled by the full
density-matrix model below rather than by an embedded two-level transition;
calling the reduced selective-pulse routines on a non-spin-1 site raises a clear
error.

\section{Full Density-Matrix Model}
\label{sec:nqr-full}

The \code{spin\_dynamics.nqr.full\_dynamics} module propagates the complete
\((2I+1)\)-level density matrix in a rotating frame at the pulse carrier. It is
the required model for spin-3/2 and also runs for spin-1. The static
Hamiltonian \(H_Q+H_Z\) is diagonalized for the actual EFG and field
orientation, the RF and detection operators \(\hat{\mathbf{e}}\cdot\mathbf{I}\)
are transformed into that eigenbasis, and the density matrix is propagated
through each constant-Hamiltonian interval by
\(U_k=\exp(-iH_k\,\Delta t_k)\), \(\rho_{k+1}=U_k\rho_k U_k^{\dagger}\).

Here \code{nutation\_hz} is the bare field nutation \(\gamma B_1/(2\pi)\), so the
realized Rabi rate on a transition \(a\!-\!b\) is
\(2\,\nu_1\,|\langle a|\hat{\mathbf{e}}_1\cdot\mathbf{I}|b\rangle|\); the same
pulse is a different flip angle on different transitions, which is the physical
effect the full model captures. A single-carrier rotating-wave approximation is
used, valid when one carrier addresses one transition band (the spin-3/2
zero-field and weak-Zeeman regime).

\begin{lstlisting}[language=Python]
import numpy as np
from spin_dynamics.nqr import QuadrupolarSite, simulate_full_fid

site = QuadrupolarSite(spin=1.5, isotope="35Cl",
                       quadrupole_frequency_hz=30e6, eta=0.1)

fid = simulate_full_fid(
    site,
    nutation_hz=20e3,            # bare field nutation gamma*B1/(2*pi)
    pulse_duration_seconds=10e-6,
    times_seconds=np.linspace(0.0, 200e-6, 1024),
)
print(fid.signal)               # complex baseband signal
\end{lstlisting}

A two-pulse (Hahn-style) echo is available through \code{simulate\_full\_echo},
and an optional \code{relaxation=NQRRelaxationModel(...)} adds Liouville-space
\(T_1/T_2\) decay. The lower-level primitives \code{pulse\_hamiltonian},
\code{static\_hamiltonian\_rotating}, and \code{detection\_operator} are exposed
for custom sequences. The full model has been validated against an exact
lab-frame propagator (the rotating-wave error scales as the
\(\Omega_1^2\) Bloch-Siegert shift), and its spin-1 powder nutation curve
reproduces the classic \(119^\circ\) maximum while the spin-3/2 curve peaks at a
slightly smaller flip angle.

\section{Model Selection}
\label{sec:nqr-model-selection}

\code{select\_nqr\_model} reads the reduced-vs-full choice from the static
Hamiltonian and the RF matrix elements for the coil polarization, not from the
spin or the line count. It builds the pulse-addressed connected set (states the
pulse can drive from the target pair, plus RF-driven near-degenerate partners)
and the isolation ratio
\(\Delta_{\mathrm{iso}}/\max(\Omega_1,1/t_p,\Gamma)\).

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import QuadrupolarSite, select_nqr_model

site = QuadrupolarSite(spin=1.5, quadrupole_frequency_hz=30e6, eta=0.1)
choice = select_nqr_model(
    site, "x",
    nutation_hz=5e3,
    pulse_duration_seconds=50e-6,
    b1_direction_pas=(1, 1, 1),
)
print(choice.recommended_model)   # "full" -- spin-3/2 Kramers doublets
print(choice.describe())          # diagnostic report
\end{lstlisting}

The reduced model is recommended only when the addressed set is a single,
non-degenerate, RF-active pair \emph{and} the isolation ratio clears
\code{isolation\_threshold} (default \(5\)). The returned
\code{NQRModelSelection} reports the target states, nearest competing RF-active
transition, isolation distance and ratio, pulse bandwidth, the addressed-state
set, a degeneracy flag, and human-readable reasons. It handles the
regime-defining cases: spin-1 \(\eta=0\) and unresolved small-\(\eta\)
splittings go full via the isolation ratio; spin-3/2 Kramers doublets go full
because the addressed set spans four states; a strongly Zeeman-resolved
spin-3/2 line can return reduced as a verified special case; and a polarization
that makes neighboring lines RF-dark keeps a crowded spectrum reducible. The
check is currently advisory---the reduced workflows do not yet invoke it
automatically.

\section{Orientations and Powder Averages}

NQR is a solid-state experiment, so the signal depends on the orientation of
the RF field relative to the local EFG principal axes. A single crystal is one
fixed orientation. A powder is represented by a normalized spherical quadrature
grid:

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import powder_average_grid, single_crystal_orientation

single = single_crystal_orientation(alpha=0.0, beta=1.57079632679)
powder = powder_average_grid(n_theta=16, n_phi=32)
\end{lstlisting}

For weak-B0 calculations, each orientation may also specify the direction of
\(B_0\) in the EFG frame. If no separate \(B_0\) direction is supplied, the
current helper uses the local RF direction as the default field direction for
that sample.

Weak-field spectra use a correlated powder grid for the static field and RF
field directions. The helper \code{b0\_b1\_powder\_average\_grid} samples the
orientation of \(B_0\) in the EFG frame and the rotation angle of \(B_1\)
around \(B_0\), preserving a chosen lab-frame angle between the two fields.
For a conventional transverse coil one uses \(B_1\perp B_0\).

\section{Weak Static B0 Spectra}

Weak static magnetic fields are modeled by exact diagonalization of
\[
  H = H_Q + H_Z, \qquad
  H_Z = -2\pi\gamma\,\mathbf{B}_0\cdot\mathbf{I}.
\]
The helper is intended for the NQR regime
\[
  \epsilon_Z = \frac{|\gamma B_0|}{\nu_{\mathrm{ref}}} \ll 1,
\]
and reports the largest perturbation ratio in the result. It can warn or raise
if the requested field is no longer weak compared with the selected NQR line.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import simulate_weak_b0_spectrum

weak = simulate_weak_b0_spectrum(
    chlorine,
    b0_tesla=1e-3,
    orientations="powder",
    broadening_hz=200.0,
    weak_ratio_threshold=0.05,
)

print(weak.max_perturbation_ratio)
\end{lstlisting}

For each crystallite, the code diagonalizes \(H_Q+H_Z\), keeps transitions
near the selected zero-field NQR line, and weights them by the RF projection
\[
  I_{ab} \propto w_k
  \left|\hat{\mathbf{b}}_{1,k}\cdot\mathbf{d}_{ab,k}\right|^2,
\]
where \(w_k\) is the orientation weight. This RF projection is important for
spin-3/2 in weak fields: the eigenvalue splitting can contain several nearby
transitions, but transitions dark to the coil should not contribute to the
plotted spectrum. In the axial limit \(\eta=0\), \(B_0\parallel z_{\mathrm{PAS}}\),
and \(B_1\perp B_0\), a spin-3/2 site gives the expected pair
\(\nu_Q\pm\gamma B_0\).

\section{SLSE Detection}

The spin-lock spin-echo (SLSE) sequence is the classic pulsed NQR detection
sequence. The helper models it as repeated selective pulses on a spin-1
detection transition and reports echo amplitudes at the requested echo spacing.
Echo decay can be represented by a scalar \(T_{2e}\) envelope or by a
phenomenological Liouville-space relaxation model with \(T_1\) and \(T_2\).

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import simulate_slse, slse_sequence

sequence = slse_sequence(
    "x",
    pulse_duration_seconds=25e-6,
    nutation_hz=10e3,
    echo_spacing_seconds=1e-3,
    num_echoes=16,
)

result = simulate_slse(
    site,
    sequence,
    orientations="powder",
    t2e_seconds=20e-3,
)
\end{lstlisting}

The returned \code{SLSEResult} includes echo times, powder-averaged echo
amplitudes, per-orientation echo amplitudes, orientation weights, transition
metadata, and the eigensystem used for the first orientation. When a relaxation
model is supplied, the result also reports local cycle eigenvalues and
cycle-derived effective decay times.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import NQRRelaxationModel

relaxed = simulate_slse(
    site,
    sequence,
    orientations="powder",
    relaxation=NQRRelaxationModel(t1_seconds=1.0, t2_seconds=20e-3),
)
\end{lstlisting}

Offset and pulse-period sweeps are available as compact diagnostics for the
SLSE modulation:

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import simulate_slse_offset_sweep

sweep = simulate_slse_offset_sweep(
    site,
    "x",
    offsets_hz=[-2e3, 0.0, 2e3],
    pulse_duration_seconds=25e-6,
    nutation_hz=10e3,
    echo_spacing_seconds=500e-6,
    relaxation=NQRRelaxationModel(t2_seconds=20e-3),
)
\end{lstlisting}

\section{EFG Inhomogeneity and SLSE Spectra}

Static EFG disorder is modeled as an isochromat ensemble of independent
quadrupolar sites. Each isochromat has its own \(\nu_Q\), \(\eta\), transition
frequency, and normalized weight. This covers impurity-induced broadening,
static temperature gradients, and other inhomogeneous mechanisms analogous to
NMR isochromat broadening.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import (
    SelectivePulse,
    gaussian_efg_distribution,
    simulate_fid_efg_distribution,
)
import numpy as np

distribution = gaussian_efg_distribution(
    site,
    quadrupole_std_hz=2e3,
    samples=41,
)

fid = simulate_fid_efg_distribution(
    distribution,
    "x",
    times_seconds=np.linspace(0.0, 20e-3, 512),
    excitation=SelectivePulse("x", duration_seconds=2.5e-6, nutation_hz=100e3),
)
\end{lstlisting}

The distribution simulators check whether a coarse EFG grid can artificially
rephase within the simulated duration. Increase the number of isochromats when
the warning appears, or pass \code{rephase\_action="ignore"} only after a
convergence check.

In an SLSE experiment, the spectrum is not the FFT of the echo train. It is the
FFT of the averaged echo acquired over a receiver window \(T_{\mathrm{acq}}\)
centered on a selected echo, with \(T_{\mathrm{acq}}\) shorter than the pulse
spacing. The finite acquisition window itself broadens the spectrum:

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import simulate_slse_acquisition_spectrum

slse_spectrum = simulate_slse_acquisition_spectrum(
    distribution,
    sequence,
    acquisition_duration_seconds=200e-6,
    acquisition_points=256,
    echo_index=-1,
    noise={"target_snr": 20.0, "seed": 123},
    deconvolution_strength=1e-2,
)
\end{lstlisting}

Noise is added to the complex time-domain acquisition before the FFT. The
optional deconvolution performs a regularized inversion of the same finite-window
FFT operator. It is useful for exploring acquisition-window broadening, but it
can amplify noise and should be checked across plausible SNR values.

\section{Two-Frequency Population Transfer}

The two-frequency NQR workflow follows the perturbation-plus-detection idea
used in 2D NQR. A selective pulse on one transition changes populations in the
shared three-level spin-1 manifold; a later SLSE block detects the changed
amplitude on another transition. The normalized diagnostic is
\[
  \Delta(p,q) = \frac{S(p,q)}{S_0(q)} - 1,
\]
where \(p\) is the perturbation transition and \(q\) is the detection
transition.

\begin{lstlisting}[language=Python]
from spin_dynamics.nqr import SelectivePulse, simulate_population_transfer

transfer = simulate_population_transfer(
    site,
    SelectivePulse("x", duration_seconds=50e-6, nutation_hz=10e3),
    sequence,
    orientations="powder",
)

print(transfer.normalized_difference)
\end{lstlisting}

The diagnostic plotting example constructs a compact \(3\times3\) map over the
spin-1 \code{x}, \code{y}, and \code{z} transitions. Off-diagonal features
indicate that perturbation and detection frequencies are connected through the
same quadrupolar site.

\chapter{First Workflows}

\section{CPMG}

Application code should normally start with the public workflow runners:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_tuned_cpmg

result = run_tuned_cpmg(numpts=101, maxoffs=10)
print(result.del_w.shape)
print(result.echo.shape)
print(result.snr)
\end{lstlisting}

The asymptotic CPMG runners return \code{CPMGResult} objects with the offset
grid, asymptotic magnetization, received spectrum, time-domain echo, acquisition
time vector, optional SNR, and probe label.

\section{Finite Echo Trains}

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_matched_cpmg_train

train = run_matched_cpmg_train(
    numpts=51,
    num_echoes=8,
    auto_refine_grid=True,
    num_workers=None,
)
\end{lstlisting}

Finite train results contain one acquired spectrum and one echo per echo index.
They also include trapezoidal echo integrals and physical echo-center times.
For long trains, use the rephasing checks or enable
\code{auto\_refine\_grid=True} so the offset grid is refined before RF matrices
are built.

\subsection{CPMG-IR Trains}

Finite inversion-recovery trains add an inversion delay axis before the CPMG
readout:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_matched_cpmg_ir_train

ir = run_matched_cpmg_ir_train(
    numpts=31,
    num_echoes=4,
    tauvect=[0.5e-3, 2e-3, 6e-3, 12e-3],
    auto_refine_grid=True,
)
\end{lstlisting}

The returned arrays are indexed by inversion delay first and echo number
second. This makes the workflow useful for compact \(T_1\)-encoded echo-train
tests without manually assembling the inversion pulse, delay, excitation, and
refocusing blocks.

\subsection{Probe Parameter Sweeps}

Probe-aware CPMG workflows can also be swept over resonator \(Q\) or mistuning:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_tuned_q_sweep

sweep = run_tuned_q_sweep(q_values=[20, 50, 100], numpts=51)
\end{lstlisting}

Use the finite-train sweep variants when echo-to-echo evolution matters. The
finite sweep results keep one \code{CPMGTrainResult} per parameter value and
also stack echo integrals for quick heat-map style summaries.

\section{FID}

\begin{lstlisting}[language=Python]
from spin_dynamics.parameters import set_params_ideal_fid
from spin_dynamics.workflows.fid import sim_fid_ideal

sp, pp = set_params_ideal_fid(numpts=101)
macq, fid, tvect = sim_fid_ideal(sp, pp)
\end{lstlisting}

\section{Imaging}

\begin{lstlisting}[language=Python]
import numpy as np
from spin_dynamics.workflows import (
    make_imaging_field_maps,
    run_tuned_phase_encoded_cpmg_imaging,
    form_imaging_image,
)

rho = np.eye(8)
fields = make_imaging_field_maps(
    rho,
    b0_map=np.zeros_like(rho),
    b1_tx_map=np.ones_like(rho),
    b1_rx_map=np.ones_like(rho),
)
result = run_tuned_phase_encoded_cpmg_imaging(
    fields,
    num_echoes=4,
    ny=9,
    receive_mode="weighted",
)
image = form_imaging_image(result, mode="echo_sum")
\end{lstlisting}

Imaging workflows return raw k-space, reconstructed per-echo images, field
maps, relaxation maps, gradient metadata, and sequence timing. Image formation
is intentionally separated from acquisition so selected-echo, echo-summed,
fitted-\(\rho\), and fitted-\(T_2\) displays can be compared.
\code{make\_imaging\_field\_maps} accepts spatial B0, transmit-B1, and
receive-B1 maps, and the ideal imaging path can include an inversion-recovery
preparation through \code{run\_t1\_encoded\_phase\_encoded\_cpmg\_imaging}.
When \code{noise} is supplied, clean k-space and image fields remain in the
result while noisy counterparts are stored in \code{kspace\_noisy},
\code{image\_noisy}, and \code{magnitude\_noisy}.

\section{Received-Signal Noise}

Noise is opt-in and output-referred. High-level workflows accept a
\code{NoiseSpec}, a mapping, or a scalar white-noise standard deviation:

\begin{lstlisting}[language=Python]
from spin_dynamics.noise import NoiseSpec
from spin_dynamics.workflows import run_tuned_cpmg

noisy = run_tuned_cpmg(
    numpts=101,
    noise=NoiseSpec(model="probe", target_snr=25.0, seed=3),
)
\end{lstlisting}

\code{model="white"} adds flat complex receiver noise. \code{model="probe"}
uses the probe output noise density where the workflow can supply it, which is
important when the received spectrum is filtered by a tuned, untuned, or
matched resonator. Result objects preserve clean deterministic fields and add
noisy fields plus \code{NoiseMetadata} so repeated SNR studies remain
reproducible.

\section{Diffusion}

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_matched_diffusion_q_sweep

sweep = run_matched_diffusion_q_sweep(
    q_values=[20, 50],
    num_echoes=3,
    numpts=21,
    dz=50e-6,
    auto_refine_grid=True,
)
\end{lstlisting}

The compact matched-probe diffusion workflow is solver-validated through
\(Q=2000\). Higher-Q cases remain a stiffness target for the current NumPy
matched-probe transient solver.

Matched diffusion CPMG also accepts \code{absolute\_phase}.  In this combined
path, the diffusion-encoding \(\pi\) pulse and the CPMG refocusing pulses are
solved at their laboratory-frame RF phase, while the free-precession intervals
retain the diffusion attenuation:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_matched_diffusion_cpmg

combined = run_matched_diffusion_cpmg(
    num_echoes=16,
    echo_spacing_seconds=1.0e-3,
    diffusion_time=1.0e-3,
    q_value=50,
    absolute_phase={
        "rf_frequency_hz": 0.25 / 1.0e-3,
        "phase_bins": 16,
    },
)

metadata = combined.absolute_phase
encoding_phase = metadata.encoding_absolute_phase_rad
refocus_phases = metadata.refocus_absolute_phase_rad
\end{lstlisting}

\code{metadata.refocus\_*} describes the echo-train refocusing pulses.
\code{metadata.encoding\_*} describes the diffusion-encoding \(\pi\) pulse,
and \code{pulse\_*} fields provide the full RF event list.

The comparison plotting example toggles diffusion and absolute phase
independently:

\begin{lstlisting}[language=bash]
python examples\plot_diffusion_absolute_phase_compare.py
python examples\plot_tuned_diffusion_absolute_phase_compare.py
\end{lstlisting}

The diffusion plotting examples use a narrow default \code{--dz-um} and enable
automatic offset-grid refinement so the displayed decay is not an artifact of
discrete-grid rephasing.  They print the effective number of offsets after any
refinement.  Use \code{--no-auto-refine-grid} only when intentionally testing
the rephasing guard.
The matched diffusion comparison also prints the matched-probe absolute-phase
residual and warns when it is below the plotting tolerance.  The current
matched pulse-shape solver is often nearly phase-invariant; in that case the
plot is a diagnostic limitation rather than evidence for a measurable combined
effect.
The tuned-probe comparison uses the same diffusion kernel with tuned
pulse-shape solves and usually shows a much larger residual, making it the
better contrast example for phase-sensitive probe transients.

\subsection{PGSE and PGSE-Prepared CPMG}

For Stejskal-Tanner pulsed-gradient diffusion encoding, use the PGSE workflow
helpers. The deterministic backend tracks the effective gradient moment and
therefore evaluates
\[
  b = \int q(t)^2\,\dd t,
  \qquad
  q(t)=\gamma\int_0^t G_{\mathrm{eff}}(t')\,\dd t' .
\]
For a rectangular PGSE pair this reduces to
\[
  b = (\gamma G \delta)^2\left(\Delta-\frac{\delta}{3}\right),
\]
where \(\delta\) is the lobe duration and \(\Delta\) is the leading-edge
separation between the two gradient lobes. The physical lobes are scheduled
with the same polarity around the refocusing pulse; the 180-degree pulse flips
the coherence-frame sign, so stationary spins refocus while diffusive motion
attenuates the signal.

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_pgse_moment

pgse = run_pgse_moment(
    num_echoes=8,
    gradient_amplitude=0.12,      # T/m
    gradient_duration=2.5e-3,     # delta
    diffusion_time=28.0e-3,       # Delta
    diffusion_coefficient=1.2e-9,
    first_echo_time_seconds=56e-3,
    echo_spacing_seconds=8e-3,
    t2_seconds=80e-3,
)
\end{lstlisting}

\code{num\_echoes=1} is the ordinary spin-echo PGSE acquisition. Larger echo
counts represent a PGSE preparation followed by a compact CPMG acquisition.
For spatially explicit studies, \code{run\_pgse\_walkers} uses the
moving-isochromat sequence driver and seeded Brownian walkers. It is slower and
stochastic, but can be extended to inhomogeneous field maps, boundaries, and
advection:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_pgse_walkers

walkers = run_pgse_walkers(
    gradient_amplitude=0.05,
    gradient_duration=2.0e-3,
    diffusion_time=16.0e-3,
    diffusion_coefficient=2.3e-9,
    walkers_per_cell=12000,
    seed=123,
)
\end{lstlisting}

The PGSE tests compare the moment backend to the analytical
Stejskal-Tanner formula, and compare the walker backend against the same
attenuation through a zero-diffusion walker reference.

\subsection{Restricted Diffusion and Diffusive Diffraction}

Because the walker backend integrates explicit displacements, it can model
diffusion restricted by hard walls---physics the closed-form moment backend
cannot express. The \code{boundary} argument of \code{run\_pgse\_walkers}
accepts the rectangular-box modes \code{"reflect"}, \code{"periodic"}, and
\code{"clip"}, or any callable mapping particle positions to confined
positions. The helper \code{make\_circular\_reflector(center, radius)} supplies
a reflecting circular wall for a pore.

For a slab pore of width \(L\) along the gradient axis, confining the walkers
makes the echo attenuation bend above the free Stejskal-Tanner line
\(E=\exp(-bD)\): once the diffusion length \(\sqrt{2D\Delta}\) approaches \(L\)
the spins can no longer fully dephase, and the apparent diffusion coefficient
\(D_{\mathrm{app}}=-\ln(E)/b\) falls below the bulk value \(D\) as the diffusion
time grows. The example
\code{plot\_pgse\_restricted\_diffusion.py} sweeps several pore widths and
diffusion times to show both signatures.

In a closed two-dimensional pore the restricted signal becomes non-monotonic.
In the narrow-pulse, long-mixing limit (\(\delta\ll a^2/D\ll\Delta\)) the
normalized echo approaches the squared magnitude of the pore form factor, the
\emph{diffusive diffraction} regime of \(q\)-space NMR. For a uniform disc of
radius \(a\),
\[
  E(q)\;\longrightarrow\;\left|\frac{2\,J_1(q_{\mathrm{ang}}a)}{q_{\mathrm{ang}}a}\right|^2,
\]
with minima at the zeros of \(J_1\), i.e. \(q_{\mathrm{ang}}a=3.83,\,7.02,\,
10.17,\dots\). Note the factor-of-\(2\pi\) ambiguity in the definition of
\(q\): the \emph{angular} wavenumber is \(q_{\mathrm{ang}}=\gamma G\delta\)
(units \(\mathrm{rad/m}\), with \(b=q_{\mathrm{ang}}^2(\Delta-\delta/3)\)),
while the \emph{reciprocal-space} (Callaghan) convention uses
\(q=\gamma G\delta/(2\pi)\) (units \(1/\mathrm{m}\), encoding kernel
\(\exp(i2\pi q x)\)), for which the first disc feature sits near \(qa\sim1\).

\begin{lstlisting}[language=Python]
import numpy as np
from spin_dynamics.motion import make_circular_reflector
from spin_dynamics.workflows import run_pgse_walkers

radius = 5.0e-6
axis = np.linspace(-radius, radius, 21)
xx, zz = np.meshgrid(axis, axis, indexing="ij")
rho = (xx**2 + zz**2 <= radius**2).astype(float)  # uniform disc

walkers = run_pgse_walkers(
    rho=rho, x_axis=axis, z_axis=axis,
    gradient_amplitude=2.0,        # large G reaches the q-space regime
    gradient_duration=0.4e-3,      # short delta (narrow-pulse limit)
    diffusion_time=80.0e-3,        # long Delta >> a^2 / D
    diffusion_coefficient=2.3e-9,
    boundary=make_circular_reflector((0.0, 0.0), radius),
    walkers_per_cell=28, substeps_per_interval=80, seed=2026,
)
\end{lstlisting}

The example \code{plot\_pgse\_circular\_pore\_diffraction.py} plots the
resulting fringes against the Bessel form factor. It offers two methods. The
default \code{walker-sweep} re-runs the confined PGSE simulation at every
gradient and so captures finite-pulse blurring of the minima. The optional
\code{--method sgp-propagator} exploits the fact that the walker trajectories
are independent of \(q\): it runs the confined diffusion once, records each
walker's net displacement \(\Delta x\) over the diffusion time, and evaluates
\(E(q)=\bigl|\langle\exp(iq_{\mathrm{ang}}\Delta x)\rangle\bigr|\) for all
\(q\) in a single vectorized sum. This is the average-propagator (\(q\)-space)
method; it is tens of times faster and reproduces the same diffraction pattern
in the \(\delta\to0\) limit. Keep the per-substep hop \(\sqrt{2D\,\dd t}\) well
below the pore size so reflection stays accurate.

\section{Moving Isochromats}

The motion layer treats inhomogeneity in the same spirit as static
isochromats, but the samples move through two-dimensional B0, transmit-B1, and
receive-B1 maps:

\begin{lstlisting}[language=Python]
from spin_dynamics.motion import make_motion_field_maps_2d, initialize_ensemble_from_density
from spin_dynamics.sequences.motion import run_motion_cpmg_sequence

fields = make_motion_field_maps_2d([-1, 0, 1], [-1, 0, 1])
ensemble = initialize_ensemble_from_density([[0, 1, 0], [1, 2, 1], [0, 1, 0]])
motion = run_motion_cpmg_sequence(
    ensemble,
    fields,
    num_echoes=4,
    echo_spacing=1e-3,
    excitation_duration=20e-6,
    refocusing_duration=40e-6,
    gradient=(0.0, 0.1),
)
\end{lstlisting}

Lower-level helpers such as \code{free\_precession\_with\_motion\_step} are
useful for custom trajectories. The sequence helpers split long intervals into
substeps so field maps are sampled along the particle path rather than only at
the beginning of an interval.

\section{Time-Varying Fields}

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import (
    sinusoidal_field_waveform,
    run_ideal_time_varying_amplitude_sweep,
)

waveform = sinusoidal_field_waveform(num_echoes=16)
sweep = run_ideal_time_varying_amplitude_sweep(
    amplitudes=[0, 0.5, 1.0, 2.0],
    waveform=waveform,
    numpts=101,
    auto_refine_grid=True,
)
\end{lstlisting}

\section{WURST}

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import (
    run_ideal_wurst_inversion,
    run_matched_wurst_cpmg,
)

ideal = run_ideal_wurst_inversion(numpts=101)
matched = run_matched_wurst_cpmg(num_echoes=2, numpts=51, num_steps=64)
\end{lstlisting}

\section{Absolute RF Phase}

Finite CPMG train workflows accept an optional \code{absolute\_phase}
mapping.  The ideal workflow records the laboratory-frame phase schedule while
preserving its nominal rectangular pulses.  Probe-aware finite trains use the
same schedule to solve the tuned, un-tuned, or matched probe waveform for each
pulse, discretize the rotating-frame waveform into short segments, and build a
separate refocusing matrix for each echo:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_tuned_cpmg_train

train = run_tuned_cpmg_train(
    numpts=21,
    num_echoes=32,
    absolute_phase={
        "rf_frequency_hz": 0.25 / 200e-6,
        "rf_phase_at_zero_rad": 0.0,
    },
)

phase_step = train.absolute_phase.delta_refocus_phase_cycles
\end{lstlisting}

For long trains, add \code{phase\_bins} to reuse solved refocusing pulses on a
uniform absolute-phase grid:

\begin{lstlisting}[language=Python]
train = run_tuned_cpmg_train(
    num_echoes=256,
    absolute_phase={
        "rf_frequency_hz": 0.03125 / 200e-6,
        "phase_bins": 32,
    },
)

library = train.absolute_phase.refocus_pulse_library
matrix_phases = train.absolute_phase.refocus_matrix_phase_rad
\end{lstlisting}

The scheduled phase for every echo remains available as
\code{refocus\_absolute\_phase\_rad}.  The quantized matrix phase, bin index,
unique phase list, and exported \code{PulseShapeLibrary} are recorded in result
metadata so cache behavior can be audited or reused in downstream analysis.
The matched diffusion CPMG workflow uses the same \code{absolute\_phase}
mapping and additionally records the diffusion-encoding \(\pi\) pulse in
\code{encoding\_*} metadata fields.

The same solved pulse shapes can be inspected without running a full echo
train:

\begin{lstlisting}[language=Python]
import numpy as np

from spin_dynamics.pulse_diagnostics import solve_probe_pulse_shape_sweep

sweep = solve_probe_pulse_shape_sweep(
    probe="tuned",
    absolute_phase_rad=2 * np.pi * np.array([0.0, 0.25, 0.5, 0.75]),
    numpts=17,
)

shape = sweep.shapes[1]
drive = shape.drive
quadrature_fraction = shape.quadrature_energy_fraction
library = sweep.pulse_shape_library
\end{lstlisting}

When sweeping the phase advance between refocusing pulses, keep the absolute
phase of the first refocusing pulse fixed if the goal is to isolate
echo-to-echo variation.  A half-cycle phase advance produces the same
transverse rotating-frame pulse shape for a linear tuned probe, while
intermediate advances change the rise/fall and quadrature transients and can
produce extra attenuation in the matched-filter echo amplitude.

The Mandal-2015 plotting examples expose the diagnostic workflow directly:

\begin{lstlisting}[language=bash]
python examples\plot_mandal2015_phase_step_sweep.py --probe tuned
python examples\plot_mandal2015_echo_modulation.py --probe tuned
python examples\plot_mandal2015_pulse_shapes.py --probe tuned
\end{lstlisting}

The finite-train plotting scripts accept \code{--phase-bins N} to exercise the
same quantized pulse-cache path used by the workflow API.
They also enable automatic offset-grid refinement by default and use
\code{--rephase-action raise}; this keeps demonstration plots from silently
using too few isochromats.  Pass \code{--no-auto-refine-grid} only for explicit
diagnostic runs.

\code{plot\_mandal2015\_pulse\_shapes.py} solves a single refocusing pulse at
several absolute RF phases and plots \(\mathrm{Re}(B_1)\),
\(\mathrm{Im}(B_1)\), \(|B_1|\), and the phase of the nonzero-amplitude
segments.  It is intended as a debugging and teaching aid: the finite train
examples use the same public diagnostics API, pulse-shape solvers, and
discretized segments.

\section{Radiation Damping}

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_radiation_damping_fid

fid = run_radiation_damping_fid(
    probe="matched",
    fill_factor=0.7,
    equilibrium_magnetization=0.8,
    flip_angle=1.0,
    model="circuit",
    detuning=2.0e4,
)
\end{lstlisting}

Finite tuned and matched CPMG trains accept an opt-in
\code{radiation\_damping} mapping:

\begin{lstlisting}[language=Python]
from spin_dynamics.workflows import run_tuned_cpmg_train

train = run_tuned_cpmg_train(
    numpts=51,
    num_echoes=8,
    radiation_damping={
        "fill_factor": 0.7,
        "equilibrium_magnetization": 0.8,
        "model": "circuit",
        "detuning": 2.0e4,
        "apply_during_pulses": True,
    },
)
\end{lstlisting}

The CPMG train path is deterministic and keeps the radiation-damping feedback
inside the finite acquisition model. For strongly inverted samples,
\code{simulate\_nmr\_maser} uses the same feedback machinery with a pumped
longitudinal magnetization; it is intended as an idealized threshold and
saturation diagnostic rather than a full spectrometer-control model.

\section{Inverse Laplace Analysis}

\begin{lstlisting}[language=Python]
import numpy as np
from spin_dynamics.analysis import invert_t2, t2_kernel

echo_times = np.linspace(0.5e-3, 90e-3, 40)
t2_axis = np.logspace(-4, -1, 60)
signal = t2_kernel(echo_times, t2_axis) @ np.exp(-((np.log(t2_axis) + 4.5) ** 2))

result = invert_t2(
    signal,
    echo_times,
    t2_axis,
    regularization=5e-4,
    regularization_order=2,
)
\end{lstlisting}

For two-dimensional distributions, the separable forward model is
\[
  D = K_1 F K_2^{\mathsf{T}},
\]
where \(D\) is the measured data, \(F\) is the unknown distribution, and
\(K_1\), \(K_2\) are relaxation or diffusion kernels. Use
\code{invert\_t1\_t2} or \code{invert\_d\_t2} for common NMR cases.

The PGSE D-T2 example ties this analysis layer to a sequence model: it builds
the b-axis with \code{run\_pgse\_moment}, simulates a PGSE-prepared CPMG echo
matrix, and recovers the D-T2 distribution with the SciPy-backed non-negative
\code{invert\_d\_t2} solver.

\section{Optimization Workflows}

The optimization package provides compact, plotting-free drivers for
phase-program searches and result handoff:

\begin{lstlisting}[language=Python]
from spin_dynamics.optimization import run_tuned_refocusing_multistart

opt = run_tuned_refocusing_multistart(
    num_segments=3,
    num_starts=4,
    numpts=21,
    max_passes=4,
)
\end{lstlisting}

Use the optimization examples as diagnostics rather than turnkey production
pulse design: they expose refocusing, target-excitation, and inverse-excitation
objective behavior while the inverse-cancellation workflow remains under
validation.

\chapter{Examples}

The example scripts can be run from \code{PythonSpinDynamics}; each script also
works from inside \code{examples/} by adding the local source tree to
\code{sys.path} when needed.

\begin{lstlisting}[language=bash]
python examples\ideal_cpmg.py --numpts 101
python examples\ideal_fid.py --numpts 101
python examples\ideal_cpmg_train.py --numpts 101 --num-echoes 8
python examples\ideal_time_varying_cpmg.py --numpts 101 --num-echoes 16
python examples\compare_cpmg_fid.py --numpts 101
python examples\export_validation_arrays.py results\validation_arrays.npz --numpts 101
python examples\tuned_probe_cpmg.py --numpts 101
python examples\probe_cpmg_compare.py --numpts 101
python examples\probe_parameter_sweeps.py --numpts 101
python examples\matched_cpmg_ir_train.py --numpts 21 --num-echoes 4 --num-tau 4
python examples\finite_probe_train_sweeps.py --numpts 21 --num-echoes 3
python examples\matched_diffusion_cpmg.py --numpts 21 --num-echoes 3
python examples\received_signal_noise.py --numpts 51
python examples\radiation_damping_fid.py --probe matched --points 401
python examples\radiation_damping_cpmg_train.py --probe tuned --numpts 21 --num-echoes 4
python examples\nmr_maser.py
python examples\heteronuclear_j_editing.py --points 33
python examples\coupled_isochromat_fields.py --points 21
python examples\diagnose_optimization_backends.py --backend all --numpts 21 --segments 3
\end{lstlisting}

Plotting examples require Matplotlib:

\begin{lstlisting}[language=bash]
python examples\plot_ideal_workflows.py --numpts 201 --output results\ideal_workflows.png
python examples\plot_ideal_imaging.py --pixels 6 --ny 7 --output results\ideal_imaging.png
python examples\plot_custom_imaging_fields.py --pixels 8 --ny 7 --output results\custom_imaging_fields.png
python examples\plot_inverse_laplace.py --output results\inverse_laplace.png
python examples\plot_pgse_d_t2.py --output results\pgse_d_t2.png
python examples\plot_probe_parameter_sweep.py --probe tuned --sweep q --output results\tuned_q_sweep.png
python examples\plot_probe_cpmg.py --numpts 101 --output results\probe_cpmg.png
python examples\plot_finite_train_workflows.py --numpts 65 --num-echoes 4 --output results\finite_trains.png
python examples\plot_diffusion_sweep.py --numpts 65 --num-echoes 3 --output results\diffusion_sweep.png
python examples\plot_diffusion_absolute_phase_compare.py --output results\diffusion_absolute_phase_compare.png
python examples\plot_tuned_diffusion_absolute_phase_compare.py --output results\tuned_diffusion_absolute_phase_compare.png
python examples\plot_time_varying_sweep.py --numpts 51 --num-echoes 12 --output results\time_varying_sweep.png
python examples\plot_motion_linear.py --output results\motion_linear.png
python examples\plot_motion_diffusion_cpmg.py --output results\motion_diffusion_cpmg.png
python examples\plot_radiation_damping.py --output results\radiation_damping.png
python examples\plot_radiation_damping_detuning.py --output results\rd_detuning.png
python examples\plot_radiation_damping_cpmg_train.py --output results\rd_cpmg.png
python examples\plot_mandal2015_phase_step_sweep.py --probe tuned --output results\mandal_phase_sweep.png
python examples\plot_mandal2015_echo_modulation.py --probe tuned --output results\mandal_echo_modulation.png
python examples\plot_mandal2015_pulse_shapes.py --probe tuned --output results\mandal_pulse_shapes.png
python examples\plot_nmr_maser.py --output results\nmr_maser.png
python examples\plot_wurst_flow.py --numpts 41 --num-steps 64 --output results\wurst_flow.png
python examples\plot_j_editing_spectrum.py --output results\j_editing_spectrum.png
python examples\plot_j_editing_field_spread.py --output results\j_editing_field_spread.png
python examples\plot_tango_filter.py --target 160 --output results\tango_filter.png
python examples\plot_slic_two_spin.py --j-hz 7 --delta-hz 0.7 --output results\slic_two_spin.png
python examples\plot_optimization_workflows.py --numpts 11 --segments 2 --starts 2 --inverse-starts 4 --output results\optimization_workflows.png
python examples\plot_optimization_pipeline.py --numpts 11 --refocusing-segments 2 --refocusing-starts 2 --excitation-segments 2 --excitation-starts 2 --inverse-starts 3 --output results\optimization_pipeline.png
python examples\plot_nqr_powder_nutation.py --output results\nqr_powder_nutation.png
python examples\plot_nqr_full_powder_nutation.py --output results\nqr_full_powder_nutation.png
python examples\plot_nqr_auto_model_selection.py --output results\nqr_auto_model_selection.png
python examples\plot_nqr_population_transfer.py --output results\nqr_population_transfer.png
python examples\plot_nqr_slse_offset.py --output results\nqr_slse_offset.png
python examples\plot_nqr_slse_spacing.py --output results\nqr_slse_spacing.png
python examples\plot_nqr_efg_broadening.py --output results\nqr_efg_broadening.png
python examples\plot_nqr_temperature_broadening.py --output results\nqr_temperature_broadening.png
python examples\plot_nqr_slse_efg_broadening.py --output results\nqr_slse_efg_broadening.png
python examples\plot_nqr_weak_b0_spectrum.py --output results\nqr_weak_b0_spectrum.png
\end{lstlisting}

\chapter{Validation}

Validation fixtures live in \code{validation/fixtures}. The tests compare the
Python implementation against compact MATLAB or Octave arrays for core
rotations, echo conversion, FID, the \code{arb10} kernel, finite acquisition,
probe response, imaging, pulse-shape helpers, optimization result handling, and
public workflow result shapes.

Regenerate dependency-light fixtures with Octave or MATLAB from the
\code{PythonSpinDynamics} directory:

\begin{lstlisting}[language=bash]
matlab -batch "run('validation/octave/generate_basic_fixtures.m')"
matlab -batch "run('validation/octave/generate_imaging_fixtures.m')"
matlab -batch "run('validation/octave/generate_optimization_fixtures.m')"
matlab -batch "run('validation/octave/generate_pulse_fixtures.m')"
\end{lstlisting}

Matched-probe fixtures require MATLAB toolboxes used by the reference
implementation. Octave fixture generation skips matched-probe files when
\code{fmincon} is unavailable.

\chapter{API Reference}

This chapter documents the intended public surface. Lower-level module imports
remain available for validation and porting. For application code, start with:

\begin{lstlisting}
spin_dynamics.workflows
spin_dynamics.parameters
spin_dynamics.analysis
spin_dynamics.coupling
spin_dynamics.nqr
spin_dynamics.noise
spin_dynamics.motion
spin_dynamics.sequences.motion
spin_dynamics.optimization
\end{lstlisting}

Drop into lower-level modules only when a building block is needed directly.
Use \code{spin\_dynamics.coupling} for the scoped dense scalar-coupled
spin-1/2 tools described in Chapter~\ref{chap:jcoupling}. Use
\code{spin\_dynamics.nqr} for the selective pulsed NQR tools described in
Chapter~\ref{chap:nqr}.

\section{Workflow Results}

\begin{longtable}{p{0.43\linewidth}p{0.49\linewidth}}
\toprule
Class & Purpose \\
\midrule
\code{CPMGResult} & Asymptotic CPMG spectrum, echo, time vector, SNR, and probe label. \\
\code{CPMGTrainResult} & Finite echo-train spectra, echoes, echo integrals, and sequence timing. \\
\code{CPMGIRTrainResult} & Inversion-recovery finite trains over a \code{tauvect}. \\
\code{MatchedCPMGIRTrainResult} & Matched-probe specialization of the CPMG-IR train result. \\
\code{CPMGParameterSweepResult} & Asymptotic Q or mistuning sweeps. \\
\code{CPMGFiniteParameterSweepResult} & Finite-train Q or mistuning sweeps. \\
\code{ZMagnetizationSweepResult} & Matched-probe z-magnetization Q sweeps. \\
\code{IdealTimeVaryingCPMGResult} & Final echo for ideal time-varying-field CPMG. \\
\code{ProbeTimeVaryingCPMGResult} & Probe-aware time-varying-field CPMG result. \\
\code{IdealCPMGImagingResult} & Ideal phase-encoded CPMG imaging result. \\
\code{ProbeCPMGImagingResult} & Tuned or matched phase-encoded CPMG imaging result. \\
\code{ImagingEchoFitResult} & Voxel-wise mono-exponential echo-stack fit. \\
\code{ImagingNoiseStatistics} & Repeated-trial image-domain noise summary. \\
\code{MotionSequenceResult} & Moving-isochromat sequence samples and final particle ensemble. \\
\code{NoiseMetadata} & Generated received-noise parameters and realized SNR. \\
\code{MatchedDiffusionCPMGResult} & Matched diffusion-aware CPMG train. \\
\code{MatchedDiffusionQSweepResult} & Matched diffusion Q sweep. \\
\code{PGSEMomentResult} & Deterministic PGSE or PGSE-prepared CPMG attenuation from gradient moments. \\
\code{PGSEWalkerResult} & Random-walker PGSE signal, sequence samples, and initial particle ensemble. \\
\code{WURSTInversionResult} & WURST inversion profile and pulse metadata. \\
\code{MatchedWURSTCPMGResult} & Matched WURST-CPMG echoes and spectra. \\
\code{RadiationDampingFIDResult} & Radiation-damping FID simulation and analytic comparison fields. \\
\code{MultiStartOptimizationResult} & Repeated random-start pulse-optimization result. \\
\code{RefocusingOptimizationResult} & Bounded fixed-amplitude refocusing phase optimization. \\
\code{ExcitationOptimizationResult} & Bounded fixed-amplitude target-excitation or inverse-excitation optimization. \\
\code{TunedExcitationInversePipelineResult} & Selected-refocusing to excitation/inverse-excitation pipeline result. \\
\code{SLSEResult} & NQR SLSE echo train, local orientation signals, and transition metadata. \\
\code{SLSESweepResult} & NQR SLSE selected-echo amplitudes and effective decay estimates across a sweep. \\
\code{NQRFIDDistributionResult} & EFG-distribution FID, FFT spectrum, and isochromat metadata. \\
\code{SLSEAcquisitionSpectrumResult} & Finite-window SLSE echo acquisition and spectrum, including optional noise and deconvolution metadata. \\
\code{WeakB0SpectrumResult} & Static weak-B0 transition spectrum and perturbation-ratio metadata. \\
\code{PopulationTransferResult} & NQR perturbation-plus-SLSE signal, reference signal, and normalized difference. \\
\bottomrule
\end{longtable}

\section{High-Level Workflows}

\begin{lstlisting}
from spin_dynamics.workflows import (
    run_ideal_cpmg, run_tuned_cpmg, run_untuned_cpmg, run_matched_cpmg,
    run_ideal_cpmg_train, run_tuned_cpmg_train,
    run_untuned_cpmg_train, run_matched_cpmg_train,
    run_ideal_cpmg_ir_train, run_tuned_cpmg_ir_train,
    run_untuned_cpmg_ir_train, run_matched_cpmg_ir_train,
    run_tuned_q_sweep, run_matched_q_sweep,
    run_tuned_mistuning_sweep, run_matched_mistuning_sweep,
    run_tuned_finite_q_sweep, run_untuned_finite_q_sweep,
    run_matched_finite_q_sweep,
    run_tuned_finite_mistuning_sweep,
    run_untuned_finite_mistuning_sweep,
    run_matched_finite_mistuning_sweep,
    run_matched_z_magnetization_q_sweep,
    run_matched_diffusion_cpmg, run_matched_diffusion_q_sweep,
    run_pgse, run_pgse_moment, run_pgse_walkers,
    pgse_b_value, gradient_moment_b_value,
    run_ideal_time_varying_cpmg_final,
    run_tuned_time_varying_cpmg_final,
    run_untuned_time_varying_cpmg_final,
    run_matched_time_varying_cpmg_final,
    run_ideal_time_varying_amplitude_sweep,
    run_tuned_time_varying_amplitude_sweep,
    run_untuned_time_varying_amplitude_sweep,
    run_matched_time_varying_amplitude_sweep,
    sinusoidal_field_waveform,
    run_ideal_wurst_inversion,
    run_matched_wurst_inversion,
    run_matched_wurst_cpmg,
    run_radiation_damping_fid,
)
\end{lstlisting}

The main CPMG runners share \code{numpts} and \code{maxoffs}. Finite train
runners add \code{num\_echoes}, relaxation constants, rephasing controls,
optional \code{auto\_refine\_grid}, and optional worker controls. Probe-aware
train runners also accept \code{q\_value} and \code{mistuning\_offset} where the
underlying probe model supports them.

\section{Imaging API}

\begin{lstlisting}
from spin_dynamics.workflows import (
    ImagingFieldMaps,
    make_imaging_field_maps,
    load_imaging_field_maps_npz,
    run_ideal_phase_encoded_cpmg_imaging,
    run_tuned_phase_encoded_cpmg_imaging,
    run_matched_phase_encoded_cpmg_imaging,
    run_t1_encoded_phase_encoded_cpmg_imaging,
    reconstruct_image_from_kspace,
    form_imaging_image,
    fit_imaging_echo_decay,
    summarize_imaging_noise_trials,
)
\end{lstlisting}

Prefer the \code{phase\_encoded} names for new code. The shorter compatibility
aliases remain available:

\begin{lstlisting}
run_ideal_cpmg_imaging
run_tuned_cpmg_imaging
run_matched_cpmg_imaging
\end{lstlisting}

\section{Parameter Constructors}

\begin{lstlisting}
from spin_dynamics.parameters import (
    set_params_ideal,
    set_params_ideal_fid,
    set_params_tuned_orig,
    set_params_untuned_orig,
    set_params_matched_orig,
    set_params_tuned_spa,
    set_params_untuned_spa,
    set_params_matched_spa,
    set_params_tuned_jmr,
    set_params_untuned_jmr,
    set_params_matched_jmr,
)
\end{lstlisting}

These constructors return dataclass instances mirroring MATLAB \code{sp},
\code{pp}, and \code{params} structures. They accept optional \code{numpts}
arguments for lightweight tests and examples while preserving MATLAB defaults
when omitted.

\section{Analysis API}

\begin{lstlisting}
from spin_dynamics.analysis import (
    t2_kernel, t1_kernel, diffusion_kernel, laplace_kernel,
    invert_laplace_1d, invert_laplace_2d,
    invert_t2, invert_t1, invert_t1_t2, invert_d_t2,
    default_regularization_strengths,
    estimate_noise_rms_from_snr,
    expected_residual_norm_from_snr,
    select_regularization_1d,
    select_regularization_2d,
)
\end{lstlisting}

For one-dimensional relaxation distributions, use:

\begin{lstlisting}
invert_t2
invert_t1
\end{lstlisting}

For separable two-dimensional maps, use:

\begin{lstlisting}
invert_t1_t2
invert_d_t2
\end{lstlisting}

The \code{select\_regularization\_*} helpers choose Tikhonov strengths from an
RMS SNR estimate.

\section{Optimization Diagnostics}

Optimization helpers expose compact NumPy/SciPy-compatible diagnostics for
phase-program searches:

\begin{lstlisting}[language=Python]
from spin_dynamics.optimization import run_tuned_refocusing_multistart

result = run_tuned_refocusing_multistart(num_segments=3, num_starts=4)
\end{lstlisting}

The plotting-free analysis helpers convert MATLAB-style result cells into
score summaries, and the tuned excitation/inverse pipeline carries a selected
refocusing candidate into bounded target-excitation and inverse-cancellation
tests. The inverse stage is still best treated as a parity and diagnostics
surface while strong inverse-pulse cancellation remains under validation.

\section{Core Numerical API}

\begin{lstlisting}
from spin_dynamics.core.echo import (
    calc_time_domain_echo,
    calc_time_domain_echo_arb,
    calc_fid_time_domain,
)
from spin_dynamics.core.rotations import (
    calc_rotation_matrix,
    calc_rot_axis_arba3,
    calc_rot_axis_arba4,
    calc_v0crit,
    sim_spin_dynamics_asymp_mag3,
    sim_spin_dynamics_exc,
)
from spin_dynamics.core.kernels import (
    sim_spin_dynamics_arb7,
    sim_spin_dynamics_arb10,
    sim_spin_dynamics_arb10_chunked,
    sim_spin_dynamics_arb10_diffusion,
    sim_spin_dynamics_arb10_radiation_damping,
)
from spin_dynamics.core.isochromats import (
    analyze_rephasing,
    check_rephasing,
    estimate_rephase_time,
    recommended_numpts_for_rephasing,
)
\end{lstlisting}

Core functions are intended for validation, debugging, and continued porting.
They are closer to MATLAB's array contracts than the workflow layer.

\section{Probe and Pulse API}

\begin{lstlisting}
from spin_dynamics.probes.tuned import (
    tuned_probe_lp,
    tuned_probe_rx,
    tuned_probe_rx_tf,
    calc_rot_axis_tuned_probe_lp_orig2,
    calc_masy_tuned_probe_lp_orig,
)
from spin_dynamics.probes.untuned import (
    untuned_probe_lp,
    untuned_probe_rx,
    untuned_probe_rx_tf,
    calc_rot_axis_untuned_probe_lp,
    calc_masy_untuned_probe_lp,
)
from spin_dynamics.probes.matched import (
    matching_network_design2,
    find_coil_current,
    find_coil_current_wurst,
    matched_probe_rx,
    calc_rot_axis_matched_probe,
    calc_masy_matched_probe_orig,
    calc_masy_matched_nut,
)
from spin_dynamics.pulses import (
    quantize_phase,
    create_wurst_pulse,
    adjust_untuned_segment_lengths,
    tuned_rectangular_pulse_response,
    untuned_rectangular_pulse_response,
    matched_rectangular_pulse_response,
    matched_wurst_pulse_response,
)
\end{lstlisting}

\section{Noise, Motion, and Radiation Damping}

\begin{lstlisting}
from spin_dynamics.noise import (
    NoiseSpec,
    NoiseMetadata,
    MatchedFilterSNRResult,
    add_received_noise,
    estimate_matched_filter_snr,
    ideal_noise_density,
    tuned_probe_output_noise_density,
    untuned_probe_output_noise_density,
    matched_probe_output_noise_density,
)
from spin_dynamics.motion import (
    ParticleEnsemble,
    MotionFieldMaps2D,
    make_motion_field_maps_2d,
    initialize_ensemble_from_density,
    advect_diffuse_positions,
    move_ensemble,
    apply_free_precession,
    apply_rf_rotation,
    free_precession_with_motion_step,
    receive_signal,
    transverse_b1_magnitude,
)
from spin_dynamics.sequences.motion import (
    MotionSequenceStep,
    MotionSequenceResult,
    make_motion_cpmg_sequence,
    run_motion_sequence,
    run_motion_cpmg_sequence,
)
from spin_dynamics.radiation_damping import (
    radiation_damping_time,
    proton_thermal_magnetization_density,
    water_proton_sample,
    hyperpolarized_proton_sample,
    normalized_radiation_damping_weights,
    radiation_damping_probe_from_tuned,
    radiation_damping_probe_from_matched,
    initial_state_from_flip_angle,
    analytic_radiation_damping_envelope,
    simulate_radiation_damping,
    simulate_radiation_damping_fid,
    simulate_nmr_maser,
)
\end{lstlisting}

\section{Scalar Coupling API}

\begin{lstlisting}
from spin_dynamics.coupling import (
    CoupledSpinSystem,
    CoupledIsochromatEnsemble,
    CoupledIsochromatStep,
    CoupledIsochromatSequenceResult,
    coupled_spin_system,
    coupled_isochromat_ensemble,
    free_precession_step,
    rf_step,
    simulate_coupled_isochromat_sequence,
    spin_operator,
    total_operator,
    product_operator,
    zeeman_hamiltonian,
    secular_j_hamiltonian,
    isotropic_j_hamiltonian,
    rf_hamiltonian,
    equilibrium_density,
    evolve_density,
    expectation,
    j_modulation_curve,
    proton_detected_j_modulation,
    carbon_detected_j_modulation,
    tango_b_filter,
    fit_known_j_spectrum,
    JEditingFitResult,
    simulate_slic_spectrum,
    two_spin_slic_transfer_time,
    SLICSpectrumResult,
)
\end{lstlisting}

\begin{longtable}{p{0.42\linewidth}p{0.50\linewidth}}
\toprule
Object & Purpose \\
\midrule
\code{CoupledSpinSystem} & Validated small spin-1/2 system with offsets and scalar couplings in hertz. \\
\code{CoupledIsochromatEnsemble} & Ensemble of local B0/B1 field samples for one coupled-spin system. \\
\code{free\_precession\_step}, \code{rf\_step} & Sequence-step builders with optional time-varying B0/B1 overrides. \\
\code{simulate\_coupled\_isochromat\_sequence} & Dense coupled-spin sequence propagation over an isochromat ensemble. \\
\code{spin\_operator}, \code{total\_operator}, \code{product\_operator} & Dense spin-1/2 operator builders. \\
\code{zeeman\_hamiltonian} & Rotating-frame offset Hamiltonian in radians per second. \\
\code{secular\_j\_hamiltonian} & Weak-coupling \(I_zI_z\) scalar Hamiltonian. \\
\code{isotropic\_j\_hamiltonian} & Isotropic scalar Hamiltonian for strongly coupled spin-1/2 systems. \\
\code{rf\_hamiltonian} & RF nutation Hamiltonian for selected spins. \\
\code{equilibrium\_density}, \code{evolve\_density}, \code{expectation} & Minimal dense density-matrix utilities. \\
\code{j\_modulation\_curve} & Sum of cosine J-modulation components. \\
\code{proton\_detected\_j\_modulation} & Proton-detected low-field J-editing response. \\
\code{carbon\_detected\_j\_modulation} & Carbon-detected \(\cos^n\) J-editing response for \(\mathrm{CH}_n\) groups. \\
\code{tango\_b\_filter} & Ideal TANGO-B scalar-coupling filter envelope. \\
\code{fit\_known\_j\_spectrum} & Least-squares amplitudes for supplied J-coupling positions. \\
\code{simulate\_slic\_spectrum} & Spin-lock response scan for dense coupled-spin systems. \\
\bottomrule
\end{longtable}

\section{ESR API}

\begin{lstlisting}
from spin_dynamics.esr import (
    ESRSpinSystem,
    ESRRelaxationModel,
    ESRDistributionSample,
    NuclearSite,
    ElectronNuclearSystem,
    resonance_frequency_hz,
    resonance_field_tesla,
    effective_g_value,
    simulate_field_sweep,
    simulate_frequency_spectrum,
    static_disorder_grid,
    simulate_field_sweep_distribution,
    simulate_frequency_spectrum_distribution,
    gaussian_lineshape,
    lorentzian_lineshape,
    flip_angle_duration,
    simulate_fid,
    simulate_hahn_echo,
    electron_nuclear_system,
    diagonalize_hyperfine_system,
    simulate_hyperfine_field_sweep,
)
\end{lstlisting}

\begin{longtable}{p{0.42\linewidth}p{0.50\linewidth}}
\toprule
Object & Purpose \\
\midrule
\code{ESRSpinSystem} & Single-electron spin-1/2 ESR system with scalar, diagonal, or full \(g\) tensor. \\
\code{resonance\_frequency\_hz}, \code{resonance\_field\_tesla} & Convert between field and microwave frequency for one ESR orientation. \\
\code{effective\_g\_value} & Direction-dependent \(g_{\mathrm{eff}}=|\mathbf{g}^{\mathsf{T}}\hat{\mathbf{n}}|\). \\
\code{simulate\_field\_sweep}, \code{simulate\_frequency\_spectrum} & Single-crystal or powder CW ESR spectra with Gaussian/Lorentzian absorption or derivative display. \\
\code{gaussian\_lineshape}, \code{lorentzian\_lineshape} & Unit-height absorption profiles and first derivatives. \\
\code{static\_disorder\_grid} & Weighted diagonal \(g\)-strain and applied-field-offset samples. \\
\code{simulate\_field\_sweep\_distribution} & Field-swept ESR spectrum averaged over static disorder samples. \\
\code{flip\_angle\_duration} & Rectangular-pulse duration for a spin-1/2 flip angle and nutation rate. \\
\code{simulate\_fid}, \code{simulate\_hahn\_echo} & Rotating-frame pulsed ESR FID and Hahn-echo density-matrix simulations. \\
\code{ESRRelaxationModel} & Phenomenological Liouville-space \(T_1/T_2\) relaxation model for pulsed ESR. \\
\code{NuclearSite}, \code{ElectronNuclearSystem} & Dense electron-nuclear product-space hyperfine model containers. \\
\code{electron\_nuclear\_system} & Convenience builder for isotropic electron-nuclear hyperfine systems. \\
\code{diagonalize\_hyperfine\_system} & Exact dense energy levels and ESR-active transitions for a hyperfine system. \\
\code{simulate\_hyperfine\_field\_sweep} & Field-swept ESR spectrum including isotropic electron-nuclear hyperfine coupling. \\
\bottomrule
\end{longtable}

\section{NQR API}

\begin{lstlisting}
from spin_dynamics.nqr import (
    QuadrupolarSite,
    NQRTransition,
    NQREigensystem,
    spin_matrices,
    quadrupole_frequency_scale_hz,
    quadrupole_hamiltonian,
    zeeman_hamiltonian,
    nqr_hamiltonian,
    diagonalize_site,
    OrientationSample,
    single_crystal_orientation,
    powder_average_grid,
    b0_b1_powder_average_grid,
    b0_powder_average_grid,
    SelectivePulse,
    apply_selective_pulse,
    transition_drive_scale,
    NQRRelaxationModel,
    slse_sequence,
    simulate_slse,
    simulate_slse_offset_sweep,
    simulate_slse_spacing_sweep,
    gaussian_efg_distribution,
    temperature_efg_distribution,
    simulate_fid_efg_distribution,
    simulate_slse_acquisition_spectrum,
    simulate_weak_b0_spectrum,
    weak_field_ratio,
    zeeman_frequency_hz,
    simulate_population_transfer,
    SLSEResult,
    PopulationTransferResult,
    WeakB0SpectrumResult,
)
\end{lstlisting}

\begin{longtable}{p{0.42\linewidth}p{0.50\linewidth}}
\toprule
Object & Purpose \\
\midrule
\code{QuadrupolarSite} & One quadrupolar nucleus in the EFG principal-axis system. \\
\code{spin\_matrices} & Dense angular-momentum operators for arbitrary integer or half-integer spin. \\
\code{quadrupole\_frequency\_scale\_hz} & Spin-dependent Hamiltonian scale for spin-1 and spin-3/2 frequency conventions. \\
\code{quadrupole\_hamiltonian} & Zero-field quadrupole Hamiltonian in radians per second. \\
\code{zeeman\_hamiltonian} & Optional weak-B0 perturbation in radians per second. \\
\code{diagonalize\_site} & Energy levels, transition frequencies, and transition dipole vectors. \\
\code{OrientationSample} & One EFG orientation relative to lab RF and optional B0 fields. \\
\code{single\_crystal\_orientation} & One fixed orientation sample. \\
\code{powder\_average\_grid} & Normalized spherical powder quadrature grid. \\
\code{b0\_b1\_powder\_average\_grid} & Powder grid with correlated static-field and RF-field directions. \\
\code{SelectivePulse} & Rectangular selective pulse on one transition. \\
\code{apply\_selective\_pulse} & Embedded two-level RF propagation in the energy basis. \\
\code{NQRRelaxationModel} & Phenomenological Liouville-space population/coherence relaxation rates. \\
\code{slse\_sequence} & Convenience builder for SLSE detection. \\
\code{simulate\_slse} & Selective-pulse SLSE echo train for a fixed orientation or powder. \\
\code{simulate\_slse\_offset\_sweep}, \code{simulate\_slse\_spacing\_sweep} & SLSE diagnostics versus RF offset or pulse period. \\
\code{gaussian\_efg\_distribution}, \code{temperature\_efg\_distribution} & Static EFG isochromat distributions. \\
\code{simulate\_fid\_efg\_distribution} & Time-domain FID and FFT spectrum from an EFG distribution. \\
\code{simulate\_slse\_acquisition\_spectrum} & Finite-window acquired SLSE echo and spectrum, with noise and optional deconvolution. \\
\code{simulate\_weak\_b0\_spectrum} & Static weak-B0 transition spectrum for spin-1 or spin-3/2. \\
\code{weak\_field\_ratio}, \code{zeeman\_frequency\_hz} & Weak-field regime checks based on \(|\gamma B_0|/\nu_{\mathrm{ref}}\). \\
\code{simulate\_population\_transfer} & Perturbation pulse plus SLSE detection for 2D NQR diagnostics. \\
\bottomrule
\end{longtable}

\section{Optimization API}

\begin{lstlisting}
from spin_dynamics.optimization import (
    spa_pulse_list,
    rectangular_refocusing_lengths,
    evaluate_tuned_refocusing_pulse,
    evaluate_untuned_refocusing_pulse,
    evaluate_matched_refocusing_pulse,
    evaluate_ideal_v0crit_refocusing_pulse,
    evaluate_ideal_v0crit_excited_refocusing_pulse,
    evaluate_ideal_time_varying_refocusing_pulse,
    optimize_tuned_refocusing_phases,
    optimize_untuned_refocusing_phases,
    optimize_matched_refocusing_phases,
    optimize_ideal_v0crit_refocusing_phases,
    optimize_ideal_v0crit_excited_refocusing_phases,
    optimize_ideal_time_varying_refocusing_phases,
    run_tuned_refocusing_multistart,
    run_untuned_refocusing_multistart,
    run_matched_refocusing_multistart,
    run_ideal_v0crit_refocusing_multistart,
    run_ideal_v0crit_excited_refocusing_multistart,
    run_ideal_time_varying_refocusing_multistart,
    evaluate_tuned_excitation_pulse,
    evaluate_tuned_inverse_excitation_pulse,
    optimize_tuned_excitation_phases,
    optimize_tuned_inverse_excitation_phases,
    run_tuned_excitation_multistart,
    run_tuned_inverse_excitation_multistart,
    run_tuned_excitation_inverse_pipeline,
    multistart_to_matlab_results,
    save_multistart_results_npz,
    load_optimization_results,
    summarize_matlab_results,
    select_matlab_result_program,
    analyze_matlab_optimization_results,
    analyze_tuned_inverse_result_pair,
)
\end{lstlisting}

The optimization layer is useful for compact pulse-design studies and
MATLAB-style result inspection. Broad historical \code{fmincon} parity and
strong inverse-excitation cancellation parity remain validation targets.

\chapter{Known Gaps}

The Python port is intentionally conservative where MATLAB parity still
matters. The largest remaining gaps are broader high-Q matched-probe validation,
full historical MATLAB \code{.mat} structure parity, broad \code{fmincon}
parity, paper-level scalar-coupling fixture comparisons, relaxation-aware
coupled-spin pulse-sequence runners, full nonselective quadrupolar pulse
propagation, experimental NQR fixture comparisons, ESR anisotropic hyperfine
and multi-electron interactions, and acceleration backends beyond the current
NumPy implementation.

New work should add small reference fixtures first, then a NumPy implementation,
then optional acceleration or plotting once numerical behavior is pinned down.

\end{document}
